# GLMM's in R {#GLMMsR}

When completing an analysis, I may skip around in my script quite a bit. Therefore, I like all packages loaded, and all data uploaded and manipulated at the start of my script.

```{r setup, include=FALSE, message=F, warning=FALSE, results = "hide", echo=F}
knitr::opts_chunk$set(cache = T, message =F, warning = F, echo=F,
                      # dpi = 36, out.width="600px", out.height="600px")
  fig.width = 8,
  fig.height = 5,
  out.width = '100%')

# R libraries:
# install.packages("kableExtra") ## uncomment to download any required packages
library(MASS)
library(lme4)
library(magrittr)
library(tidyverse)
library(ggmap)
# usethis::edit_r_environ()
register_google(Sys.getenv("ggmapKey"))

# to run Bayesian models. Set appropriate nCores for your machine
library(brms)
nCores = 6 # nCores -2
library(rjags)
library(kableExtra)

## to use jags, you first need to download. For a Mac, go to terminal:
# brew install jags
# instructions from: https://gist.github.com/casallas/8411082

## Set colors:
# library(RColorBrewer)
# display.brewer.pal(11, "Spectral")
# brewer.pal(11, "Spectral")
mycolors = c("#9E0142", "#66C2A5")
# mycolors =  c("#9E0142", "#762A83","#9970AB",  # "#5E4FA2",
#               "#3288BD","#66C2A5", "#ABDDA4", "#D53E4F")
visualcols = c( "#D7191C", "#FDAE61", "#FFFFBF", "#A6D96A", "#1A9641")
trialcols = c("#4575B4", "#40004B","#762A83","#9970AB",  "#5AAE61",
               "#C2A5CF","#74ADD1", "#E7D4E8", "#ABD9E9",   "#00441B" )
prop_colors = c("#313695", "#ABD9E9", "#FFFFBF", "#FEE090", 
                 "#FDAE61", "#F46D43","#D73027", "#A50026")
#
```
``` {r functions}
# calculate trapping reduction for frequentist model fits.
getTR <- function (mod, coefs = -1) {
  b1 = summary(mod)$coef[coefs,1]
  seb1 = summary(mod)$coef[coefs,2]
  (TR = round(100 * (1- exp(b1)),1))
  (higher = round(100 * (1 - exp(b1 - 1.96*seb1)),1))
  (lower = round(100 * (1 - exp(b1 + 1.96*seb1)),1)) 
  data.frame(#Treatment = names(TR),
             TR = as.numeric(TR), 
             lowerCI = as.numeric(lower),
             upperCI = as.numeric(higher))   
}

## NA = 0 in sum
`%+%` <- function(x, y)  mapply(sum, x, y, MoreArgs = list(na.rm = TRUE))
#

## prediction plot for frequentist models:
predplot <- function(preddata, plottitle = "") {
  p = ggplot(preddata,
       aes(DATI, preds, color = TreatmentF)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowerCI, ymax = upperCI, fill = TreatmentF),
              alpha = 0.3) + 
  geom_point(data = mean_cts,
             aes(DATI, mean_mothsperday, color = TreatmentF), size = 1.1) +
  facet_wrap(~Location, scales = "free_y") + 
  scale_color_manual(values = mycolors[]) +
  scale_fill_manual(values = mycolors[]) +
  labs(title = plottitle, 
       subtitle = "Lines are predictions, points are the data")

  return(p)
}

brmspredplot = function(preddata, plottitle = "") {
  p <- ggplot(preddata, 
       aes(x = DATI, y = Estimate, color = TreatmentF)) +  
  geom_line() + 
  geom_ribbon(aes(ymin = `Q2.5`, ymax = `Q97.5`, 
                  fill = TreatmentF),  # regression line and CI
                    alpha = 0.3) +
  geom_point(data = mean_cts, 
             aes(DATI, mean_mothsperday, color = TreatmentF), 
             size = 1.1) +
  facet_wrap(~Location, scales = "free_y") + 
  scale_color_manual(values = mycolors[]) +
  scale_fill_manual(values = mycolors[]) +
  labs(title = plottitle,
       subtitle = "Lines are predictions, points are the data")

  return(p)
}

```
```{r readdatR}

datRinit = read.csv("data/moths.csv")
# drop first column, is just unnecessary indexing:
datRinit %<>% dplyr::select(-X)
# (there is a select function in MASS package that sometimes overrides dplyr)
```
```{r data_manip, results = "hide"}
# check data values: 
str(datRinit)
summary(datRinit)
lapply(datRinit, function(x) {
  if (is.character(x))
    unique(x)
})
# convert dates to dates in case further manipulation on them is needed:
datR <- datRinit %>%
  mutate(TransplantDate = ymd(TransplantDate),
         DispInstallDate = ymd(DispInstallDate),
         TrapInstallDate = ymd(TrapInstallDate),
         SamplingDate = ymd(SamplingDate)
         )  %>%
  mutate(mothsperday = nYSB / DaysOfCatch, # standardize the counts
         TreatmentF = factor(Treatment), # if we need to fix levels in non-alpha order
         LocationF = as.factor(Location),
         # DATscaled = as.numeric(scale(DAT)),
         SamplingDateC = as.character(SamplingDate),
         numericDate = as.numeric(SamplingDate),
         numericDate = numericDate - min(numericDate) + 1)

## average counts for each location, date
mean_cts <- datR %>%
  group_by(Location, Treatment, TreatmentF,
           AssessmentNumber,
           SamplingDate, DATI) %>%
  summarize(mean_cts = mean(nYSB, na.rm = T),
            mean_mothsperday = mean(mothsperday)) %>%
  ungroup()


```


## Data Exploration

### Moth counts

Here is what the data look like. Columns: Location, Treatment, TrapID, Longitude, Latitude, AssessmentNumber, and SamplingDate should be self-explanatory... TransplantDate is the date that the rice seedlings were transplanted into the field; DispInstallDate is the treatment installation date; TrapInstallDate is the trap installation date; "nYSB" are the moth counts (number of YSB moths); DaysOfCatch is the number of days since the trap was previously sampled  or the number of days since trap installation if it is assessment number 1; DATI is days after trap installation; DADI is days after dispenser installation; and DAT is days after transplant. Some of these columns will be discussed more later.

```{r}
# knitr::kable(head(datRinit, 20))

# check data values:
summary(datRinit)
lapply(datRinit, function(x) {
  if (is.character(x))
    unique(x)
})
# datR <- datRinit %>%
#   mutate(TransplantDate = ymd(TransplantDate),
#          DispInstallDate = ymd(DispInstallDate),
#          TrapInstallDate = ymd(TrapInstallDate),
#          SamplingDate = ymd(SamplingDate)
#          )  %>%
#   mutate(mothsperday = nYSB / DaysOfCatch,
#          TreatmentF = factor(Treatment), # if we need to fix levels in non-alpha order
#          LocationF = as.factor(Location),
#          logDaysOfCatch = log(DaysOfCatch),
#          # DATscaled = as.numeric(scale(DAT)),
#          SamplingDateC = as.character(SamplingDate),
#          numericDate = as.numeric(SamplingDate),
#          numericDate = numericDate - min(numericDate) + 1)


```

The data are the moth counts collected at each trap (average per day), within each location and throughout the season. Note how the y-axis changes for each plot in the figure.

```{r plotMoths, fig.width = 8, fig.height = 6}
ggplot(datR,
       aes(DATI, mothsperday, color = TreatmentF)) +
  geom_jitter(height = 0, width = 0.75, alpha = 0.5) +
  # geom_line(data = mean_cts,
      # aes(DATI, mean_mothsperday, color = TreatmentF), size = 1.1) +
  geom_smooth(se=F) +
  facet_wrap(~Location, scales = "free_y") + 
  scale_color_manual(values = mycolors[]) +
  # scale_y_sqrt() +
  labs(title = "Male YSB moth counts for each location and trap",
       y = "Moth counts per trap per day",
       x = "Days after trap installation (DATI)",
       color = "Treatment")
#
```

### Trap locations

Trial locations in relation to each other. Some trial locations are closer to each other than others.

```{r}

basemap <- get_googlemap(c(lon = mean(datR$Longitude, na.rm = T),
                           lat = mean(datR$Latitude, na.rm = T)),
                         color = "bw",
                         zoom = 7)
ggmap(basemap) +
  geom_point(data = datR,
             aes(Longitude, Latitude, color = Location), size = 2) +
  labs(title = "Trial locations") +
  scale_color_brewer(palette = "PuOr") 
#
```

Within each location, the treatment traps have a slightly different alignment:

```{r}
coords <- datR %>%
  dplyr::select(Location, Treatment, TrapID, Latitude, Longitude) %>%
  unique()
ggplot(coords,
       aes(Longitude, Latitude, color = Treatment)) +
  geom_point() +
  facet_wrap(~Location, scales = "free") +
  theme(aspect.ratio=1, 
        axis.text.x=element_blank(),
        axis.text.y=element_blank()) +
  ggrepel::geom_text_repel(aes(label = TrapID), color = "black", size = 3) +
  scale_x_continuous(expand = c(0.3, 0)) + 
  scale_y_continuous(expand = c(0.3, 0)) + 
  labs(title = "Trap ID's at each location")

```


### Timeline

```{r timeline, results="hide", fig.width = 7}

dates <- datR %>%
  group_by(Location, TransplantDate, TrapInstallDate) %>%
  summarize(mindate = min(SamplingDate, na.rm = T),
            maxdate = max(SamplingDate, na.rm = T)) %>%
  ungroup() %>%
  arrange(desc(mindate), Location)
dates$Location <- factor(dates$Location, levels = unique(dates$Location))

ggplot(dates) +
  geom_errorbarh(aes(xmin = mindate, xmax = maxdate, 
                     y = Location, color = Location),
                 height = 0.5, linewidth = 1.25) +
  geom_jitter(aes(TransplantDate, y = Location), width = 0, height = 0.1, size = 1.25) +
  geom_jitter(aes(TrapInstallDate, y = Location), width = 0, height = 0.1, 
              size = 1.25, color = gray(0.6)) +
  labs(y = "Location",
       x = "Date range of moths data collection (2023)",
       title = "Trapping date ranges of each location",
       caption = "Black points are transplant dates \n
       Grey points are trap installation dates") +
  scale_color_viridis_d() 
```

Note that not all traps are installed on exactly the same day; not all locations have rice transplanted on exactly the same day; and not all traps are sampled on exactly the same day.

## Basic GLM models

To start, we ignore all the spatial and temporal relationships in the data and assume each data point is independent and identically distributed (iid). **This is not a good model for the data!** It is just our starting off point.

When building hierarchical Bayesian models, it is always a good idea to start simple, make sure everything works as expected, and then build up.

### Fitting the frequentist models

Note: when I predict for new data, I set DaysOfCatch = 1, and then I compare to the moths per day variable (mothsperday = nYSB / DaysOfCatch). I want to exclude any patterns related to the varying time intervals.

These models show a very tight confidence intervals for our predictions. But also, the residual deviance is MUCH greater than the degree of freedom, indicating a lack of fit. The plot of the predictions overlaid on the data also demonstrate the lack of fit.

```{r pois1}

mod1p <- glm(nYSB ~ TreatmentF, 
             offset = log(DaysOfCatch),
             family = poisson,
             data = datR)
summary(mod1p)
TR_1p <- getTR(mod1p)
tmp = predict(mod1p, se = T, type = "link",
              newdata = datR  %>% mutate(DaysOfCatch = 1))
# str(tmp)
preds1p <- data.frame(datR, 
                      preds = tmp$fit,
                      lowerCI = tmp$fit + 2*tmp$se.fit,
                      upperCI = tmp$fit - 2*tmp$se.fit)
preds1p %<>%
  mutate(preds = exp(preds),
         lowerCI = exp(lowerCI),
         upperCI = exp(upperCI))

p1p_glm = predplot(preds1p, "Poisson GLM predictions")
print(p1p_glm)

```

The model is such a bad fit to the data, it is hard to even tell what is going on in the plot above.

For the negative binomial model, our confidence intervals are a little wider, but we are still ignoring all the correlations in our data. And the plot of the predictions again indicates the lack of fit.

```{r nb1}
mod1nb <- glm.nb(nYSB ~ TreatmentF + 
             offset(log(DaysOfCatch)),
             data = datR)
summary(mod1nb)
TR_1nb <- getTR(mod1nb)
tmp = predict(mod1nb, se = T, type = "link",
              newdata= datR  %>% mutate(DaysOfCatch = 1))
# str(tmp)
preds1nb <- data.frame(datR, 
                      preds = tmp$fit,
                      lowerCI = tmp$fit - 2*tmp$se.fit,
                      upperCI = tmp$fit + 2*tmp$se.fit)
preds1nb %<>%
  mutate(preds = exp(preds),
         lowerCI = exp(lowerCI),
         upperCI = exp(upperCI))
# str(preds1nb)

p1nb_glm = predplot(preds1nb, "Neg. binomial GLM predictions")
print(p1nb_glm)

```

### Bayesian (brms) model fit

In R, I fit the Bayesian models using the "brms" package. I find this package very intuitive and it has fast, efficient algorithms based on Stan. 

Because the models take a few minutes to fit, I usually fit them when I am initially running through my code, save them, and then only load the model fit when rendering the Rmarkdown file and creating the resulting figures.

The prior_summary command is helpful if you do not know what a parameter is called in 'brms'. Here, I used the command to find out what they called their $\phi$ parameter. (They call it the shape parameter.)

```{r, eval = F}
brm1nb <- brm(formula = nYSB ~ TreatmentF + offset(log(DaysOfCatch)),
              data = tmp, 
              family = negbinomial, #log-link is default
              prior = c(set_prior("normal(0,2)", class = "b"),
                        set_prior("normal(0,2)", class = "Intercept"),
                        set_prior("cauchy(0,1)", class = "shape")), # is half-cauchy
              warmup = 500, iter = 2000,  # use more iter for final model
              seed = 5000, # so that you get the same results
              chains = nCores, cores = nCores) 
saveRDS(brm1nb, "../output/brm1nb.RDS")
```
```{r}
brm1nb <- readRDS("output/brm1nb.RDS")
prior_summary(brm1nb)
# 
```

The Bayesian parameter estimates match very closely to the frequentist estimates. This is expected, but reassuring that we have built the correct foundation for the more complicated models to come. The first summary output is from the Bayesian model; the second is from the frequentist fit. 

Comparisons of TR predictions and predicted moth count values can be found at the end of the page.

```{r}
summary(brm1nb)
summary(mod1nb)
```
```{r}
library(ggmcmc) # ggs()
# CALC TR!
mod1t <- ggs(brm1nb) # transform MCMC output into a table
# unique(mod1t$Parameter)
beta1 <- mod1t %>% 
  filter(Parameter == "b_TreatmenttrtA", # our beta1
         Iteration > 500) # remove burn-in
TR = 100 * (1 - exp(beta1$value))
# as_draws_df(fit$draws("b_TreatmenttrtA"))
bayesTR_1nb <- data.frame(TR = round(median(TR), 1), 
           lowerCI = round(quantile(TR, 0.025), 1),
          upperCI = round(quantile(TR, 0.975), 1))   
print(bayesTR_1nb)

# plot predictions!!!
predsbrm1nb <- fitted(brm1nb,
                       # type = "link",
                       scale = "linear",
                        newdata = datR  %>% mutate(DaysOfCatch = 1))   

bayes_p1nb = brmspredplot(cbind(datR, exp(predsbrm1nb)), 
                          "Bayesian neg. binomial GLM predictions")
print(bayes_p1nb)

```

### Bayesian (rjags) model fit

I also fit the Bayesian models using the `rjags` package. Eventually, I fit complex non-linear, hierarchical Bayesian models and for those models, I need more flexibility than `brms` provides. Because it is good practice to build up to the complex models, I am also fitting these simpler models with the rjags package. 

#### Poisson model in jags

```{r jagsSetup1p}

model_string1p <- "model{
  # Likelihood
  for(i in 1:N){
    y[i] ~ dpois(lambda[i]) 
    log(lambda[i]) <-  b0 + b1 * xA[i] + logDaysOfCatch[i]
  }
  
  # Priors
  b0 ~ dnorm(0, 1/2)
  b1 ~ dnorm(0, 1/2)
  # (normal dist uses precision not variance)

    # derived parameters:
  TR <- 100 - 100 * exp(b1)

}"

dataList = list(N= nrow(datR),
                y = datR$nYSB, 
                xA = ifelse(datR$Treatment == "CGP", 0, 1),
                logDaysOfCatch = log(datR$DaysOfCatch))

jags1p <- jags.model(file = textConnection(model_string1p), 
                    data = dataList)
```
```{r jagsRun1p}
update(jags1p, n.iter = 1000) # burn-in
idata_jags1p <- coda.samples(jags1p,
                              variable.names = c("b0", "b1", "TR"),
                              n.iter = 10000,
                              thin = 5)
write_rds(idata_jags1p, "output/jags1p.RDS")
```
```{r jagsOut1p}
idata_jags1p <- readRDS("output/jags1p.RDS")

summary(idata_jags1p)
summary(mod1p)
# check for convergence:
# plot(posterior_sample)
# heidel.diag(out)  
# raftery.diag(out)  
# effectiveSize(out)  

# plot predictions:
pars <- as.mcmc(idata_jags1p)
# str(pars)
# create a model matrix from x
X <- cbind(1, ifelse(datR$Treatment == "CGP", 0, 1))
lambda_hat <- exp(X %*% t(pars[,-1]))

predjags1p <- datR %>%
  mutate(preds = apply(lambda_hat, 1, median),
         lowerCI = apply(lambda_hat, 1, quantile, 0.025),
         upperCI = apply(lambda_hat, 1, quantile, 0.975))
predplot(predjags1p, "Bayesian Poisson GLM predictions- jags")

```

#### NB model in jags

Here I also use the matrix form of $X\beta$.

Note that jags uses a different parameterization than R., so there is an extra line of code to make them match. Also, I coded my model in matrix form here-- this parameterization of the priors is equivalent to the Poisson model above. I include both for comparison.

```{r jagsSetup1nb}
model_string1nb <- "model{
  # Likelihood
  for(i in 1:N){
    y[i] ~ dnegbin(p[i], r)
    p[i] <- r / (r+lambda[i]) 
    log(lambda[i]) <-  mu[i]
    mu[i] <- inprod(beta[],X[i,]) + logDaysOfCatch[i]

  }
  
  # Priors
  beta ~ dmnorm(muBeta,tauBeta)  # multivariate Normal prior
  r ~ dt(0, 1/2, 1)T(0,) # dunif(0,50) # switched to half Cauchy

    # derived parameters:
  TR <- 100 - 100 * exp(beta[2])
}"

dataList = list(N= nrow(datR),
                y = datR$nYSB, 
                X = cbind(1, ifelse(datR$Treatment == "CGP", 0, 1)),
                logDaysOfCatch = log(datR$DaysOfCatch),
                muBeta = rep(0, 2),
                tauBeta = diag(1/2, 2))

jags1nb <- jags.model(file = textConnection(model_string1nb), 
                    data = dataList)
```
```{r jagsRun1nb, eval = F}
update(jags1nb, n.iter = 1000) # burn-in
idata_jags1nb <- coda.samples(jags1nb,
                              variable.names = c("beta", "r", "TR"),
                              n.iter = 10000,
                              thin = 5)
write_rds(idata_jags1nb, "output/jags1nb.RDS")
```
```{r jagsOut1nb}
idata_jags1nb <- readRDS("output/jags1nb.RDS")

summary(idata_jags1nb)
# summary(mod1nb)

# check for convergence:
# plot(posterior_sample)
# heidel.diag(out)  
# raftery.diag(out)  
# effectiveSize(out)  

# plot predictions:
pars <- as.mcmc(idata_jags1nb)#[, grepl("beta", colnames(pars))]
# str(pars)
pars = pars[, grepl("beta", colnames(pars))]
# create a model matrix from x
X <- cbind(1, ifelse(datR$Treatment == "CGP", 0, 1))
lambda_hat <- exp(X %*% t(pars))

predjags1nb <- datR %>%
  mutate(preds = apply(lambda_hat, 1, median),
         lowerCI = apply(lambda_hat, 1, quantile, 0.025),
         upperCI = apply(lambda_hat, 1, quantile, 0.975))
predplot(predjags1nb, "Bayesian neg binomial GLM predictions- jags")

```

This model is not yet fully converged, so some differences from R output are expected.


## GLMM: Random effect (RE) for locations

The first fix we make to the model is acknowledging that overall average  moth pressure varies from location to location (see Fig 1 of moth counts). To make this fix, we add a location random effect (RE) and our model becomes a generalized linear mixed-effects model (GLMM or GLMER).

We also want to acknowledge that the treatment effect may vary from location to location-- sometimes we see a big difference in moth counts between control and treatment fields, and sometimes the difference is smaller. For inference though, we are only interested in the larger picture, which is the overall trapping reduction. (We are not interested in what happens at these exact locations per se, we are more interested in the average treatment effect.) Therefore, we also add a treatment random effect.

### GLMM mathematical model

I only show the Poisson version of the model, the NB version is a straightforward extension.

$$
y_{ik} \sim Pois(\lambda_{ik}) \\
log(\lambda_{ik} ) = \beta_0 + \beta_1 x_{A,i} + \gamma_{0k} x_{ik}  + \gamma_{1k} x_{A,i} x_{ik} + \mbox{ln} \left(DaysOfCatch_i\right) \\
\gamma_{0k} \sim Normal(0, \sigma^2_{0}) \\
\gamma_{1k} \sim Normal(0, \sigma^2_{1}) \\
$$

where, in addition to the variables and parameters defined for the GLM, we have

$k = 1,..., K=10$ represents the locations, 

$y_{ik}$ is moth count $i$ from location $k$,

$\lambda_{ik}$ is the expected moth count for sample $i$  from location $k$,

$\beta_0$ is the expected moth count for our control treatment for a new location,

$\beta_1$ is the expected difference in moth counts between the control group and the treatment A group for a new location, 

$\gamma_{0k}$ is the random intercept associated with location $k$, which leads to different background moth pressures at each location. All $\gamma_{0k}$ come from an iid Normal distribution.

$\gamma_{1k}$ is the random slope associated with location $k$, which leads to different treatment effects at each location. All $\gamma_{1k}$ come from an iid Normal distribution, and

$x_{ik}$ is an indicator variable that equals 1 if moth count $i$ is associated with location $k$ and 0 otherwise.

OK, technically I should be introducing matrices here because $k = 1,..., K=10$ locations, so we need a new indicator variable for each location, but I'm going to be lazy and write it like this for now.


#### Bayesian GLMM mathematical model

Here is the same model with a Bayesian framework and using a negative binomial regression:

To make our models Bayesian, we add priors to our parameters:

$$
y_{ik}\sim NegBinom(\lambda_{ik}, \phi) \\
log(\lambda_{ik} ) = \beta_0 + \beta_1 x_{A,i} + \gamma_{0k} x_{ik}  + \gamma_{1k} x_{A,i} x_{ik} + \mbox{ln} \left(DaysOfCatch_i\right) \\ 
\gamma_{0k} \sim Normal(0, \sigma^2_{0}) \\
\gamma_{1k} \sim Normal(0, \sigma^2_{1}) \\
\beta_0 \sim Normal(0, 2) \\
\beta_1 \sim Normal(0, 2) \\
\sigma^2_{0} \sim HalfCauchy(0,1) \\
\sigma^2_{1} \sim HalfCauchy(0,1) \\
\phi \sim HalfCauchy(0,1) \\
$$


#### Bayesian GLMM -- Matrix version

I DO write the Bayesian version in matrix form because it matches the coding (and is more correct given the need for a $KxK$ $Z$ design matrix):

$$
\mathbf{Y} \sim NegBinom(\boldsymbol\lambda, \phi) \\
log(\boldsymbol\lambda) = \mathbf{X}\boldsymbol\beta + \mathbf{Z}\boldsymbol\gamma  + \mbox{ln} \left(\bf{DaysOfCatch}\right) \\
\boldsymbol\beta \sim Normal(\mathbf{0}, 2\mathbf{I}) \\
\boldsymbol\gamma \sim Normal(\mathbf{0}, \boldsymbol\sigma^2 \mathbf{I}) \\
\boldsymbol\sigma^2 \sim HalfCauchy(\mathbf{0}, 1\mathbf{I})) \\
\phi \sim HalfCauchy(0,1)
$$


### GLMM frequentist fit

A couple of notes here. R always gets mad when you ask for SE's for predictions from a GLMM. Technically, you need to run simulations to get them and then they still come with an asterisk related to their reliability. (This is a reason to use the Bayesian model-- credible intervals are never based on approximations!)

When we plot our predictions, we see that we now have better estimates for the overall mean at each location, and we see how much they vary from location to location, but there is a strong temporal pattern at each location that we are missing.

```{r pois2}
mod2p <- glmer(nYSB ~ TreatmentF + (1 + TreatmentF|Location), 
             offset = log(DaysOfCatch),
             family = poisson,
             data = datR)
summary(mod2p)
TR_2p <- getTR(mod2p)
knitr::kable(TR_2p)
tmp = predict(mod2p, se = T, type = "link",
              newdata=datR %>% mutate(DaysOfCatch = 1))
# str(tmp)
preds2p <- data.frame(datR %>% mutate(DaysOfCatch = 1), 
                      preds = tmp$fit,
                      lowerCI = tmp$fit + 2*tmp$se.fit,
                      upperCI = tmp$fit - 2*tmp$se.fit)
preds2p %<>%
  mutate(preds = exp(preds),
         lowerCI = exp(lowerCI),
         upperCI = exp(upperCI))
# str(predsp2)

p2p_glmm = predplot(preds2p, "Poisson GLMM (Location) predictions")
print(p2p_glmm)
# print(p2p_glmm)

```
```{r nb2}
mod2nb <- glmer.nb(nYSB ~ TreatmentF + (1 + TreatmentF|Location), 
             offset = log(DaysOfCatch),
             data = datR)
summary(mod2nb)
TR_2nb <- getTR(mod2nb)
knitr::kable(TR_2nb)
tmp = predict(mod2nb, se = T, type = "link",
              newdata=datR %>% mutate(DaysOfCatch = 1))
# str(tmp)
preds2nb <- data.frame(datR, 
                      preds = tmp$fit,
                      lowerCI = tmp$fit + 2*tmp$se.fit,
                      upperCI = tmp$fit - 2*tmp$se.fit)
preds2nb %<>%
  mutate(preds = exp(preds),
         lowerCI = exp(lowerCI),
         upperCI = exp(upperCI))
# str(predsnb2)

p2nb_glmm = predplot(preds2nb, "Neg. binomial GLMM (Location) predictions")
print(p2nb_glmm)


```


### GLMM Bayes (brms) version

By default, the prior distributions of random slopes and intercepts are correlated in `brm`, which matches the "GLMM" frequentist fit. If the correlation is non-significant, you may want to remove that correlation to simplify your model. To remove the correlation, change the random effects term from "(1 + TreatmentF|Location)" to " (1 + TreatmentF||Location)" (has an extra vertical line), which makes them independent. 

You'll notice that the parameter estimates are slightly different from the frequentist NB model fit here-- the more complicated your model is, the more likely you will find this is true with slightly different versions of your model (here, adding priors and fitting with a different algorithm). 

```{r brmSetup2p, eval = F}
brm2nb <- brm(formula = nYSB ~ TreatmentF + offset(log(DaysOfCatch)) + 
                    (1 + TreatmentF|Location),
              data = datR, 
              family = negbinomial,
              prior = c(set_prior("normal(0,2)", class = "b"),
                        set_prior("cauchy(0,1)", class = "shape"), 
                        set_prior("normal(0,2)", class = "Intercept"),
                        set_prior("cauchy(0,1)", class = "sd")),
              control = list(adapt_delta = 0.9), # b/c of divergent warning
              warmup = 500, iter = 2000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(brm2nb, "../output/brm2nb.RDS")
```
```{r}
brm2nb <- readRDS("output/brm2nb.RDS")
# prior_summary(brm2nb) # we did not change the default prior for the RE correlation
# 
summary(brm2nb)
# summary(mod2nb)

```

```{r}
library(ggmcmc) # ggs()
# CALC TR!
modtransformed <- ggs(brm2nb) # transform MCMC output into a table
# head(modtransformed)
# summary(modtransformed)
# unique(modtransformed$Parameter)
beta1 <- modtransformed %>% 
  filter(Parameter == "b_TreatmentFtrtA", # our beta1
         Iteration > 500) # remove burn-in
TR = 100 * (1 - exp(beta1$value))
bayesTR_2nb <- data.frame(TR = round(median(TR), 1), 
           lowerCI = round(quantile(TR, 0.025), 1),
          upperCI = round(quantile(TR, 0.975), 1))   
knitr::kable(bayesTR_2nb)

# plot predictions:
predsbrm <- fitted(brm2nb,
                   scale = "linear",
                   newdata = datR  %>% mutate(logDaysOfCatch = 0))  

bayes_p2nb = brmspredplot(cbind(datR, exp(predsbrm)), 
                          "Bayesian neg. binomial GLMM (Location) predictions")
print(bayes_p2nb)


```

### GLMM Bayes (rjags) model fit

#### Poisson model in jags

Does not converge well without correlated RE's.

Also, need correlated RE's to match glmer and "brms"...
adapted from:
[https://people.ucsc.edu/~abrsvn/general_correlated_ranefs_bayes_jags.r](https://people.ucsc.edu/~abrsvn/general_correlated_ranefs_bayes_jags.r)

```{r jagsSetup2pOLD, eval = F}

# this model doesn't allow for correlation between the RE's
model_string2p <- "model{
  # Likelihood
  for(i in 1:N){
    y[i] ~ dpois(lambda[i]) 
    log(lambda[i]) <- mu[i]
    # mu[i] <- inprod(beta[], X[i,]) + gamLoc0[LocRe[i]] +  logDaysOfCatch[i]
    mu[i] <- gamLoc0[LocRe[i]] +  gamLoc1[LocRe[i]] * X[i,2] +logDaysOfCatch[i]
    #  b1 * X[i,2] +
  }
  
  # Priors
  # beta ~ dmnorm(muBeta, tauBeta)  # multivariate Normal prior
  b0 ~ dnorm(0, 1/4)
  b1 ~ dnorm(0, 1/4)
  
  sigmaLoc0 ~ dunif(0.01, 4) #dt(0, 1, 1)T(0,) # (hyperpriors)
  sigmaLoc1 ~ dunif(0.01, 4) #dt(0, 1, 1)T(0,)
  tauLoc0 <- 1 / (sigmaLoc0 * sigmaLoc0)
  tauLoc1 <- 1 / (sigmaLoc1 * sigmaLoc1)

  # priors for random effects:
  for (i in 1:NLocRE) {
    gamLoc0[i] ~ dnorm(b0, tauLoc0) # intecept-- mean is 0 (b/c of b0)
    gamLoc1[i] ~ dnorm(b1, tauLoc1) # slope-- mean is 0 (b/c of b1)
  }
  
    # derived parameters:
  TR <- 100 - 100 * exp(b1)

}"
```
```{r jagsSetup2p}

model_string2p <- "model{
  # Likelihood
  for(i in 1:N){
    y[i] ~ dpois(lambda[i]) 
    log(lambda[i]) <- mu_obs[i]
    # mu[i] <- inprod(beta[], X[i,]) + gamLoc0[LocRe[i]] +  logDaysOfCatch[i]
    mu_obs[i] <- gamLoc0[LocRe[i]] +  gamLoc1[LocRe[i]] * X[i,2] +logDaysOfCatch[i]
    #  b1 * X[i,2] +
  }
  
# Priors
# Set up the means for the multivariate ranef distribution
for (i in 1:2) {
    xi[i]~dunif(0, 100) # scaling for the multivariate RE dist (for means, sds, and the ranefs themselves)
    mu_raw[i]~dnorm(0, 1/4) # unscaled means for the multivariate RE dist
    mu[i] <- xi[i]*mu_raw[i] # scaled means for the multivariate RE dist
}
b0 <- mu[1] # mean for random intercepts
b1 <- mu[2] # mean for random slopes

# Set up the var-covar matrix for the multivariate RE dist
Tau_B_raw[1:2, 1:2] ~ dwish(W[,], 3) # W is the identity matrix, provided as data; we have 3 dofs, i.e., 2 ranefs + 1, to ensure a uniform (-1, 1) prior for the correlation between ranefs
Sigma_B_raw[1:2, 1:2] <- inverse(Tau_B_raw[,])
for (i in 1:2) {
    sigma[i] <- xi[i]*sqrt(Sigma_B_raw[i, i])
}
sigma_b0 <- sigma[1] # SD of intercepts
sigma_b1 <- sigma[2] # SD of slopes
for (i in 1:2) { for (j in 1:2) {
    rho[i, j] <- Sigma_B_raw[i, j]/sqrt(Sigma_B_raw[i, i]*Sigma_B_raw[j, j])
} }
rho_b0_b1 <- rho[1, 2]
covariance <- rho_b0_b1*sigma_b0*sigma_b1


# The multivariate ranef distribution, i.e., modeling the correlated ranefs
# priors for random effects:
for (j in 1:NLocRE) {
	B_raw_hat[j, 1] <- mu_raw[1]
	B_raw_hat[j, 2] <- mu_raw[2]
	B_raw[j, 1:2] ~ dmnorm(B_raw_hat[j, ], Tau_B_raw[, ]) # the pairs of unscaled (raw) correlated random effects
	gamLoc0[j] <- xi[1]*B_raw[j, 1] # random intercept
  gamLoc1[j] <- xi[2]*B_raw[j, 2] # random slope
}


  
  # Sampling from the prior: given that we do not place hyperpriors directly on the means, sds and correlation(s) of the multivariate ranef distribution, we want to sample from the prior to make sure we didn't accidentally make it more informed than we wanted (and we want it very vague)
for (i in 1:2) {
    xi_prior[i]~dunif(0, 100)
    mu_raw_prior[i]~dnorm(0, .0001)
    mu_prior[i] <- xi_prior[i]*mu_raw_prior[i]
}
b0_prior <- mu_prior[1]
b1_prior <- mu_prior[2]
Tau_B_raw_prior[1:2, 1:2] ~ dwish(W[,], 3)
Sigma_B_raw_prior[1:2, 1:2] <- inverse(Tau_B_raw_prior[,])
for (i in 1:2) {
    sigma_prior[i] <- xi_prior[i]*sqrt(Sigma_B_raw_prior[i, i])
}
sigma_b0_prior <- sigma_prior[1]
sigma_b1_prior <- sigma_prior[2]
for (i in 1:2) { for (j in 1:2) {
    rho_prior[i, j] <- Sigma_B_raw_prior[i, j]/sqrt(Sigma_B_raw_prior[i, i]*Sigma_B_raw_prior[j, j])
} }
rho_b0_b1_prior <- rho_prior[1, 2]
  
    # derived parameters:
  TR <- 100 - 100 * exp(b1)

}"


NLocRE = length(unique(datR$Location))
LocRe = as.numeric(as.factor(datR$Location)) # Location indexing for RE's
dataList = list(N= nrow(datR),
                NLocRE = NLocRE,
                y = datR$nYSB, 
                X = cbind(1, ifelse(datR$Treatment == "CGP", 0, 1)),
                LocRe = LocRe,
                logDaysOfCatch = log(datR$DaysOfCatch),
                W = diag(2))
# install.packages("bayesm")
library("bayesm")
var_vec <- apply(coef(mod2p)$Location, 2, var)

inits = function () {
        list(
          'xi'=rlnorm(2),
          'mu_raw'=rnorm(2), 
          'Tau_B_raw'=rwishart(3, diag(2)*var_vec)$W, 
          'sigma_res'=rlnorm(1), 
          'xi_prior'=rlnorm(2), 
          'mu_raw_prior'=rnorm(2), 
          'Tau_B_raw_prior'=rwishart(3, diag(2)*var_vec)$W
        )
}

jags2p <- jags.model(file = textConnection(model_string2p), 
                     inits = inits,
                    data = dataList, n.chains = 4)

```
```{r jagsRun2p, eval = F}
update(jags2p, n.iter = 50000) # burn-in
idata_jags2p <- coda.samples(jags2p,
                             variable.names <- c("mu", "mu_b0", "mu_b1", "sigma", "sigma_b0", "sigma_b1", "rho", "rho_b0_b1", "covariance", "b0", "b1", "sigma_res", "mu_b0_prior", "mu_b1_prior", "sigma_b0_prior", "sigma_b1_prior", "rho_b0_b1_prior",
                                                 "TR"),

                              n.iter = 10000, 
                              thin = 5)
write_rds(idata_jags2p, "output/jags2p.RDS")
```
```{r jagsOut2p, eval = F}
idata_jags2p <- readRDS("output/jags2p.RDS")

summary(idata_jags2p)[[1]]
# summary(mod2p)
getTR(mod2p)
# check for convergence:
# gelman.diag(idata_jags2p) # rhat values
# plot(idata_jags2p)
# gelman.plot(idata_jags2p)
# cor(idata_jags2p[[1]])
# effectiveSize(idata_jags2p)
# library(coda)


# plot predictions:
# pars <- as.mcmc(idata_jags2p)
pars <- runjags::combine.mcmc(idata_jags2p)
str(pars[,c("b0", "b1")])
# create a model matrix from x
X <- cbind(1, ifelse(datR$Treatment == "CGP", 0, 1))
lambda_hat <- exp(X %*% t(pars[,c("b0", "b1")]) + X %*% t())

predjags2p <- datR %>%
  mutate(preds = apply(lambda_hat, 1, median),
         lowerCI = apply(lambda_hat, 1, quantile, 0.025),
         upperCI = apply(lambda_hat, 1, quantile, 0.975))
predplot(predjags2p, "Bayesian Poisson GLMM predictions- jags")

```


**Need to add RE's for predictions!!

#### NB model in jags

Not done.


## GLMM: RE for Location, SamplingDate

In this version of the model, we acknowledge that sampling on different days of the season adds to the variability of the moth counts, and that counts from the same sampling date are more similar than for a different date. In this model, however,  the correlation between sampling dates is ignored.

This model is included here because it is important to think about whether you have nested or crossed random effects (if applicable). Here, we are assuming crossed random effects because Java island is fairly homogeneous in terms of weather and ecosystem so it may make sense for all samples from one date to be grouped together. However, one could also argue for nested RE's for this case study. 

This model allows us to include sampling date in our model as a random effect and still fit the model quickly in a frequentist framework. 

Include dates because samples within a date will be more similar than samples across all dates at a location. But only include random intercepts.

The predictions now mimic the spatial patterns we see over time.

Good nested vs crossed reference:
https://stats.stackexchange.com/questions/228800/crossed-vs-nested-random-effects-how-do-they-differ-and-how-are-they-specified

### Crossed GLMM mathematical model

I only show the Poisson version of the model, the NB version is a straightforward extension.

$$
y_{ik} \sim Pois(\lambda_{ik}) \\
log(\lambda_{ij} ) = \beta_0 + \beta_1 x_{A,i} + \gamma_{0k} x_{ik}  + \gamma_{1k} x_{A,i} x_{ik}  + \gamma_{00j} x_{ij}  + \mbox{ln} \left(DaysOfCatch_i\right) \\
\gamma_{0k} \sim Normal(0, \sigma^2_{0}) \\
\gamma_{1k} \sim Normal(0, \sigma^2_{1}) \\
\gamma_{00j} \sim Normal(0, \sigma^2_{00}) \\
$$

where, in addition to the variables and parameters defined for the GLM, we have

$k = 1,..., K=10$ represents the locations, 

$j = 1, ..., J = 117$ represents the 117 unique date-location combinations,

$y_{ikj}$ is moth count $i$ from location $k$ on date $j$,

$\lambda_{ikj}$ is the expected moth count for sample $i$  from location $k$ on date $j$,

$\beta_0$ is the expected moth count for our control treatment for a new location,

$\beta_1$ is the expected difference in moth counts between the control group and the treatment A group for a new location, 

$\gamma_{0k}$ is the random intercept associated with location $k$, which leads to different background moth pressures at each location. All $\gamma_{0k}$ come from an iid Normal distribution,

$\gamma_{1k}$ is the random slope associated with location $k$, which leads to different treatment effects at each location. All $\gamma_{1k}$ come from an iid Normal distribution, 

$x_{ik}$ is an indicator variable that equals 1 if moth count $i$ is associated with location $k$ and 0 otherwise,

$\gamma_{00j}$ is the random intercept associated with date-location $j$, which leads to different background moth pressures for each date, and 

$x_{ij}$ is an indicator variable that equals 1 if moth count $i$ is associated with date-location $j$ and 0 otherwise.


#### Bayesian GLMM -- Matrix version

$$
\mathbf{Y} \sim NegBinom(\boldsymbol\lambda, \phi) \\
log(\boldsymbol\lambda) = \mathbf{X}\boldsymbol\beta + \mathbf{Z}\boldsymbol\gamma  + \mathbf{z_{00}}\gamma_{00}  + \mbox{ln} \left(\bf{DaysOfCatch}\right) \\
\boldsymbol\beta \sim Normal(\mathbf{0}, 2\mathbf{I}) \\
\boldsymbol\gamma \sim Normal(\mathbf{0}, \boldsymbol\sigma^2 \mathbf{I}) \\
\gamma_{00} \sim Normal(0, \sigma^2) \\
\boldsymbol\sigma^2 \sim HalfCauchy(\mathbf{0}, 1\mathbf{I})) \\
\sigma^2_{00} \sim HalfCauchy(0,1) \\
\phi \sim HalfCauchy(0,1)
$$


### Crossed GLMM frequentist fit

```{r pois3, results = "hide"}
mod3p <- glmer(nYSB ~ TreatmentF  +
                 (1 + TreatmentF|Location) + 
                 (1 + TreatmentF|Location:SamplingDateC), 
             offset = log(DaysOfCatch),
             family = poisson,
             data = datR)
summary(mod3p)
TR_3p <- getTR(mod3p)
knitr::kable(TR_3p)

tmp = predict(mod3p, se = T, type = "link",
              newdata = datR %>% mutate(DaysOfCatch = 1))
# str(tmp)
preds3p <- data.frame(datR %>% mutate(DaysOfCatch = 1), 
                      preds = tmp$fit,
                      lowerCI = tmp$fit + 2*tmp$se.fit,
                      upperCI = tmp$fit - 2*tmp$se.fit)
preds3p %<>%
  mutate(preds = exp(preds),
         lowerCI = exp(lowerCI),
         upperCI = exp(upperCI))
# str(predsp3)


p3p_glmm = predplot(preds3p, "Poisson GLMM (Location, Date) predictions")
print(p3p_glmm)

```
```{r nb3, results = "hide"}
# too slow!
mod3nb <- glmer.nb(nYSB ~ TreatmentF  +
                      (1 + TreatmentF|Location) + 
                 (1 + TreatmentF|SamplingDateC), 
             offset = log(DaysOfCatch),
             data = datR)
summary(mod3nb)
TR_3nb <- getTR(mod3nb)
knitr::kable(TR_3nb)

tmp = predict(mod3nb, se = T, type = "link",
              newdata=datR %>% mutate(DaysOfCatch = 1))
# str(tmp)
preds3nb <- data.frame(datR %>% mutate(DaysOfCatch = 1), 
                       preds = tmp$fit,
                       lowerCI = tmp$fit + 2*tmp$se.fit,
                       upperCI = tmp$fit - 2*tmp$se.fit)
preds3nb %<>%
  mutate(preds = exp(preds),
         lowerCI = exp(lowerCI),
         upperCI = exp(upperCI))
# str(predsnb2)

p3nb_glmm = predplot(preds3nb, "Neg. binomial GLMM (Location, Date) predictions")
print(p3nb_glmm)

```


### Crossed GLMM Bayes (brms) version

```{r, eval = F}
brm3nb <- brm(formula = nYSB ~ TreatmentF + offset(log(DaysOfCatch)) + 
                    (1 + TreatmentF|Location) + 
                    (1|SamplingDateC),
                    # (1|Location:SamplingDateC),
              data = datR, 
              family = negbinomial,
              prior = c(set_prior("normal(0,2)", class = "b"),
                        set_prior("cauchy(0,1)", class = "shape"), # == half-cauchy
                        set_prior("normal(0,2)", class = "Intercept"),
                        set_prior("cauchy(0,1)", class = "sd")),
              warmup = 500, iter = 2000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(brm3nb, "output/brm3nb.RDS")
```
```{r}
brm3nb <- readRDS("output/brm3nb.RDS")
# prior_summary(brm3nb, all = FALSE)
# print(prior_summary(brm3nb, all = FALSE), show_df = FALSE)

summary(brm3nb)
# summary(mod3nb)
# 
```

```{r}
library(ggmcmc) # ggs()
# CALC TR!
modtranformed <- ggs(brm3nb) # transform MCMC output into a table
beta1 <- modtranformed %>% 
  filter(Parameter == "b_TreatmentFtrtA", # our beta1
         Iteration > 500) # remove burn-in
TR = 100 * (1 - exp(beta1$value))
bayesTR_3nb <- data.frame(TR = round(median(TR), 1), 
           lowerCI = round(quantile(TR, 0.025), 1),
          upperCI = round(quantile(TR, 0.975), 1))   
# knitr::kable(bayesTR_3nb)

# plot predictions:
predsbrm <- fitted(brm3nb,
                   scale = "linear",
                   newdata = datR  %>% mutate(DaysOfCatch = 1))

bayes_p3nb = brmspredplot(cbind(datR, exp(predsbrm)), 
                          "Bayesian Poisson GLMM (Location, Date) predictions")
print(bayes_p3nb)

```


## GLMM with Gaussian Processes: RE for Trial, GP for Date

Now we are in the territory where we have to fit the models in a Bayesian framework. If our response variable was continuous (i.e., the regression model was based on a Normal distribution), then we could still use maximum likelihood estimation.

### Bayesian mathematical model

$$
y_{ijk} \sim NegBinom(\lambda_{ijk}, r) \\
log(\lambda_{ijk} ) = \beta_0 + \beta_1 x_{A,i} + \gamma_{0k} x_{ik}  + \gamma_{1k} x_{A,i} x_{ik}  + GP_{ijk}  + \mbox{ln} \left(DaysOfCatch_i\right) \\
 \boldsymbol\gamma \sim Normal(\mathbf{0}, \boldsymbol\Sigma_{\gamma}) \\ 
 \mathbf{GP_k} \sim MVN(0, \boldsymbol\Sigma_k) \\
 \boldsymbol\beta \sim Normal(\mathbf{0}, 2\mathbf{I}) \\
 \boldsymbol\sigma^2 \sim HalfCauchy(\mathbf{0}, 1\mathbf{I})) \\
 r \sim HalfCauchy(0,1) \\
 \Sigma_{lm, k} = \tau \cdot e^{-\frac{|d_l - d_m|^2}{l^2}\right} \\
\tau_k \sim HalfCauchy(0, 1) \\
\l_k \sim HalfCauchy(0, 1) \\
$$

where, in addition to the variables and parameters defined for the GLM, we have

$k = 1,..., K=10$ represents the locations, 

$j = 1, ..., J = 117$ represents the 117 unique date-location combinations,

$y_{ikj}$ is moth count $i$ from location $k$ on date $j$,

$\lambda_{ikj}$ is the expected moth count for sample $i$  from location $k$ on date $j$,

$\beta_0$ is the expected moth count for our control treatment for a new location,

$\beta_1$ is the expected difference in moth counts between the control group and the treatment A group for a new location, 

$\gamma_{0k}$ is the random intercept associated with location $k$, which leads to different background moth pressures at each location. All $\gamma_{0k}$ come from an iid Normal distribution,

$\gamma_{1k}$ is the random slope associated with location $k$, which leads to different treatment effects at each location. All $\gamma_{1k}$ come from an iid Normal distribution, 

$x_{ik}$ is an indicator variable that equals 1 if moth count $i$ is associated with location $k$ and 0 otherwise,

$\gamma_{00j}$ is the random intercept associated with date-location $j$, which leads to different background moth pressures for each date, and 

$x_{ij}$ is an indicator variable that equals 1 if moth count $i$ is associated with date-location $j$ and 0 otherwise.

$l$ = lscale (differnt for each location)

$\phi$ = sdgp


### Gaussian Process Bayes (brms) version

* Use numeric version of sampling date for better algorithm stability. Still, we have a warning about a divergent transition. Ideally, we would tweak the model and look at the data carefully so fix this warning, but because we are working with made up data, we will not worry about the warning for now.

* Using squared-exponential covariance function (quite typical choice).

* Here, we have a different GP for each location (each location has its own parameters for the function). This is done to further account for expected differences between locations. Ideally, we would do some model selection to decide between this model and one that shares the GP parameters across all locations.

```{r, eval =F}
summary(datR$numericDate) # check max before fitting, in case seasons don't line up
brm4nb <- brm(formula = nYSB ~ TreatmentF + offset(log(DaysOfCatch)) + 
                    (1 + TreatmentF|Location) + 
              gp(numericDate, by =Location), 
                  control = list(adapt_delta = 0.9),
              data = datR, 
              family = negbinomial,
              prior = c(set_prior("normal(0,2)", class = "b"),
                        set_prior("cauchy(0,1)", class = "shape"), # == half-cauchy
                        set_prior("normal(0,2)", class = "Intercept"),
                        set_prior("cauchy(0,1)", class = "sd"),
                        set_prior("cauchy(0,1)", class = "lscale", 
                                  coef = "gpnumericDateLocationLoc1"),
                        set_prior("cauchy(0,1)", class = "lscale", 
                                  coef = "gpnumericDateLocationLoc2"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc3"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc4"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc5"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc6"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc7"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc8"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc9"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc10"),
                        set_prior("normal(0,2)", class = "sdgp")),
                        # set_prior("lkj(2)", class = "cor")),
              warmup = 1000, iter = 5000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(brm4nb, "output/brm4nb.RDS")
```
```{r, eval =F}
summary(datR$numericDate) # check max before fitting, in case seasons don't line up
brm4nb1gp <- brm(formula = nYSB ~ TreatmentF + offset(log(DaysOfCatch)) + 
                    (1 + TreatmentF|Location) + 
              gp(numericDate), 
                  control = list(adapt_delta = 0.9),
              data = datR, 
              family = negbinomial,
              prior = c(set_prior("normal(0,2)", class = "b"),
                        set_prior("cauchy(0,1)", class = "shape"), # == half-cauchy
                        set_prior("normal(0,2)", class = "Intercept"),
                        set_prior("cauchy(0,1)", class = "sd"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDate"),
                        set_prior("normal(0,2)", class = "sdgp")),
                        # set_prior("lkj(2)", class = "cor")),
              warmup = 1000, iter = 5000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(brm4nb1gp, "output/brm4nb1GP.RDS")
brm4nb = brm4nb1gp
```

```{r}
brm4nb <- readRDS("output/brm4nb.RDS")
summary(brm4nb)
# bayes_R2(brm4nb) # 0.824

prior_summary(brm4nb)
prior_summary(brm4nb, all = FALSE)
# 
```

```{r}
library(ggmcmc) # ggs()
# CALC TR
modtranformed <- ggs(brm4nb) # transform MCMC output into a table
beta1 <- modtranformed %>% 
  filter(Parameter == "b_TreatmentFtrtA", # our beta1
         Iteration > 500) # remove burn-in
TR = 100 * (1 - exp(beta1$value))
bayesTR_4nb <- data.frame(TR = round(median(TR), 1), 
           lowerCI = round(quantile(TR, 0.025), 1),
          upperCI = round(quantile(TR, 0.975), 1))   
knitr::kable(bayesTR_4nb)

# plot predictions:
# library(tidybayes) # add_predicted_draws()
predsbrm <- fitted(brm4nb,
                   scale = "linear",
                   newdata = datR  %>% mutate(DaysOfCatch = 1))
# dim(exp(predsbrm))
# head(predsbrm)
# I used linear scale (link response) in the fitted values and then exponeniate
# because it intuitively is easier to think about moths per day (i.e., response scale),
# but I want the resolution that is had at the linear scale.
# (If I chose response scale, Id' get lots of 0's which aren't informative. 
# Give it a try!)

bayes_p4nb = brmspredplot(cbind(datR, exp(predsbrm)), 
                          "Bayesian neg binomial GP GLMM predictions")
print(bayes_p4nb)

```

## Summary of estimates

### Compare TR estimates

Here, the trapping reduction estimates from all of the models are compared. As expected, the median TR estimates are very close from model to model, but the uncertainty in that estimate increase (i.e., the confidence/credible intervals get wider) when we properly account for the correlations in our data.

```{r compareTR}

# Framework = Frequentist, Bayes
# Distribtuion = Poisson, NB
# Model name: GLM, GLMM (Location), GLMM (Location, nested dates), GLMM (Location, GP for correlated dates)
# Model description?
# TR
# lower CI (2.5%)
# upper CI (97.5%)
tableTR = bind_rows(
data.frame(Framework = "Frequentist", Distribution = "Poisson", Model_name = "GLM",
      round(TR_1p)),
data.frame(Framework = "Frequentist", Distribution = "Neg Binomial", Model_name = "GLM",
      round(TR_1nb)),
data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLM",
      round(bayesTR_1nb)),

data.frame(Framework = "Frequentist", Distribution = "Poisson", Model_name = "GLMM (Location)",
      round(TR_2p)),
data.frame(Framework = "Frequentist", Distribution = "Neg Binomial", Model_name = "GLMM (Location)",
      round(TR_2nb)),
data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLMM (Location)",
      round(bayesTR_2nb)),

data.frame(Framework = "Frequentist", Distribution = "Poisson", Model_name = "GLMM (Location, nested Date)",
      round(TR_3p)),
data.frame(Framework = "Frequentist", Distribution = "Neg Binomial", Model_name = "GLMM (Location, nested Date)",
      round(TR_3nb)),
data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLMM (Location, nested Date)",
      round(bayesTR_3nb)),

data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLMM (Location, Gaussian Process Date)",
      round(bayesTR_4nb))
)
rownames(tableTR) <- NULL
kable(tableTR, booktabs = TRUE,
             caption = "Comparison of derived TR estimates from each model fit.") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))  %>%
  # column_spec(4, 
  #             background = "lightblue")  %>%
  pack_rows("GLM", 1, 3) %>%
  pack_rows("GLMM (Location)", 4, 6) %>%
  pack_rows("GLMM (Location, Date)", 7, 9) %>%
  pack_rows("GLMM (Location, GP)", 10, 10)

```

### Compare coefficients

Because TR is a non-linear function of $\beta_1$, the differences int he model output get slightly distorted from the transformation. Thereofre, the $\beta_1$ coefficients are displayed as well:

```{r compareBetas}

# Framework = Frequentist, Bayes
# Distribtuion = Poisson, NB
# Model name: GLM, GLMM (Location), GLMM (Location, nested dates), GLMM (Location, GP for correlated dates)
# Model description?
# TR
# lower CI (2.5%)
# upper CI (97.5%)
tableBeta = bind_rows(
data.frame(Framework = "Frequentist", Distribution = "Poisson", 
      Model_name = "GLM",
      Estimate = round(summary(mod1p)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod1p)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Frequentist", Distribution = "Neg Binomial",
           Model_name = "GLM",
      Estimate = round(summary(mod1nb)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod1nb)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", 
           Model_name = "GLM",
           Estimate = round(summary(brm1nb)$fixed["TreatmenttrtA", "Estimate"], 2),
      SE = round(summary(brm1nb)$fixed["TreatmenttrtA", "Est.Error"], 3)),


data.frame(Framework = "Frequentist", Distribution = "Poisson", 
           Model_name = "GLMM (Location)",
      Estimate = round(summary(mod2p)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod2p)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Frequentist", Distribution = "Neg Binomial",
           Model_name = "GLMM (Location)",
      Estimate = round(summary(mod2nb)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod2nb)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", 
           Model_name = "GLMM (Location)",
           Estimate = round(summary(brm2nb)$fixed["TreatmentFtrtA", "Estimate"], 2),
      SE = round(summary(brm2nb)$fixed["TreatmentFtrtA", "Est.Error"], 3)),


data.frame(Framework = "Frequentist", Distribution = "Poisson", 
           Model_name = "GLMM (Location, nested Date)",
      Estimate = round(summary(mod3p)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod3p)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Frequentist", Distribution = "Neg Binomial",
           Model_name = "GLMM (Location, nested Date)",
      Estimate = round(summary(mod3nb)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod3nb)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLMM (Location, nested Date)",
           Estimate = round(summary(brm3nb)$fixed["TreatmentFtrtA", "Estimate"], 2),
      SE = round(summary(brm3nb)$fixed["TreatmentFtrtA", "Est.Error"], 3)),

data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLMM (Location, Gaussian Process Date)",
           Estimate = round(summary(brm4nb)$fixed["TreatmentFtrtA", "Estimate"], 2),
      SE = round(summary(brm4nb)$fixed["TreatmentFtrtA", "Est.Error"], 3))
)

rownames(tableBeta) <- NULL
knitr::kable(tableBeta, booktabs = TRUE,
             caption = "Comparison of regression coefficient estimates from each model fit.")  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))  %>%
  pack_rows("GLM", 1, 3) %>%
  pack_rows("GLMM (Location)", 4, 6) %>%
  pack_rows("GLMM (Location, Date)", 7, 9) %>%
  pack_rows("GLMM (Location, GP)", 10, 10)

```



### Plot comparisons

#### GLM's

In the side-by-side comparison, you can see the wider confidence intervals for the negative binomial distribution, compared to the Poisson.

```{r all_plots1}
cowplot::plot_grid(p1p_glm + theme(legend.position="none"), 
                   p1nb_glm + theme(legend.position="none"))
```

Side-by-side comparison of the frequentist vs Bayes negative binomial GLM predictions.

```{r all_plots1b}
cowplot::plot_grid(p1nb_glm + theme(legend.position="none"),
                   bayes_p1nb + theme(legend.position="none"))
```

#### GLMM's (Location RE)

Side-by-side comparison of the frequentist Poisson vs negative  binomial GLMM (Location RE) predictions.

```{r all_plots2}
cowplot::plot_grid(p2p_glmm + theme(legend.position="none"), 
                   p2nb_glmm + theme(legend.position="none"))
```

Side-by-side comparison of the frequentist vs Bayes negative binomial GLMM (Location RE) predictions.

```{r all_plots2b}
cowplot::plot_grid(p2nb_glmm + theme(legend.position="none"),
                   bayes_p2nb + theme(legend.position="none"))
```


#### GLMM's (Location RE, nested Date)

Note: These models are a great fit to the data that we have! But, they are overly confident and will not be good at predicting at new locations and new sampling dates.

Side-by-side comparison of the frequentist Poisson vs negative  binomial GLMM (Location RE) predictions.

```{r all_plots3}
cowplot::plot_grid(p3p_glmm + theme(legend.position="none"), 
                   p3nb_glmm + theme(legend.position="none"))
```

Side-by-side comparison of the frequentist vs Bayes negative binomial GLMM (Location RE) predictions.

```{r all_plots3b}
cowplot::plot_grid(p3nb_glmm + theme(legend.position="none"),
                   bayes_p3nb + theme(legend.position="none"))
```


#### GLMM's: nested Date vs Gaussian Process Date

Once we properly account for the correlations in our sampling dates, we have much more uncertainty in our overall TR predictions (see table above) , and the predicted counts are "smoothed" over time for some of the locations.

```{r all_plots4}
cowplot::plot_grid(bayes_p3nb + theme(legend.position="none"),
                   bayes_p4nb + theme(legend.position="none"))
```

