# GLMM's in Python {#GLMMsPython}
```{r reticulate, include=FALSE, message=F, warning=FALSE, results = "hide", echo=F}

# library to use Python within R:
library(reticulate)
# py_install("pandas")
#
## make sure these packages are installed before installing pymer4:
# install.packages(c("lme4", "lmerTest", "emmeans"))

# # Ensure pymer4 is installed in the desired Python environment
# py_require("pymer4")
# 
# # Import pymer4
# pymer4 <- import("pymer4")


```
```{python, eval = F}
# You may have to install the packages if you have not used them before:
# code chunk is here for easy retrieval
# replace module name as needed.
!pip install patsy

# installing googlemaps needs a modification:
# !pip install  --use-pep517 googlemaps

# to install pymer4, you need to type the following commands from Terminal:
# I think you need Anaconda installed
## https://eshinjolly.com/pymer4/installation.html
# conda create --name pymer4 -c ejolly -c conda-forge -c defaults pymer4
# conda activate pymer4

```
```{python modules}
import pandas as pd
import pandas as pd

import bambi as bmb

import arviz as az
import matplotlib.pyplot as plt 
from formulae import design_matrices
import pymc as pm

import pickle # to write, read python objects to file
import patsy # for dmatrix function

```
```{python readdatP}

import pandas as pd

# read in the moth data:
datP = pd.read_csv("data/moths.csv", delimiter = ",")
# drop first column, is just unnecessary indexing:
datP = datP.drop(datP.columns[0], axis=1)

```
```{python data_manip, results = "hide"}

# check data values: 
datP.info()

pd.set_option('display.max_columns', None) # Display all columns
datP.describe(percentiles=[0.5], include = 'all')

for col in datP:
  if (datP[col].dtype == 'object'):
    print(col)
    print(datP[col].unique())

# convert dates to dates in case further manipulation on them is needed:
cols = ['TransplantDate', 'DispInstallDate', 'TrapInstallDate', 'SamplingDate']
for col in cols:
    datP[col] = pd.to_datetime(datP[col])
datP['SamplingDate'].describe()

# standardize the counts
datP['mothsperday'] = datP['nYSB'] / datP['DaysOfCatch']

# datR <- datP %>%
#   mutate(TreatmentF = factor(Treatment), # if we need to fix levels in non-alpha order
         # LocationF = as.factor(Location),
         
# convert sampling date to string for nested RE's:
datP['SamplingDateC'] = str(datP['SamplingDate'])
# convert sampling date to numeric starting at 1 for GP's:
datP['numericDate'] = datP['SamplingDate'].dt.dayofyear
datP['numericDate'] = datP['numericDate'] - datP['numericDate'].min() + 1
datP['numericDate'].describe()

## average counts for each location, date
mean_cts = datP.groupby(['Location', 'Treatment', 'AssessmentNumber', 'SamplingDate', 'DATI'], as_index=False).agg({'nYSB': 'mean', 'mothsperday': 'mean'})
mean_cts.describe(include='all')

```
```{python functions}
# calculate trapping reduction for frequentist model fits.
def getTR(mod):
    """Calculate TR with confidence interval for stats model"""
    # Function body
    import pandas as pd

    b1 = mod.params.iloc[1]
    seb1 = mod.bse.iloc[1]
    
    outTR = {'TR': [round(100 * 1 - math.exp(b1))], 
    'lowerCI': [round(100 * 1 - math.exp(b1 + 1.96*seb1))],
    'upperCI': [round(100 * 1 - math.exp(b1 - 1.96*seb1))]}
    df = pd.DataFrame(outTR)
    return df

```

## Getting started with Python

Python is a lot more complicated than R!

1. Download Python.

2. Decide how you want to interact with Python. Yes, you can technically use the terminal and use Python directly, but a GUI is preferable so that it is easier to run the code, add comments, view the output, etc. This is even more true for Python than it is for R. 

Originally, I downloaded Jupyter Notebook directly, but I didn't like the way things loaded. So, I downloaded "Anaconda" and then accessed Jupyter from there. However, I still don't love Jupyter-- the notebooks don't optimize the screen's size like RStudio does. In RStudio, I love having my code on one side of the screen, and my outputs on the other. Screens are wide, not long. 

Soon, I came across the "reticulate" package, which allows for Python coding in an Rmarkdown script! So far, awesome and it is how I run the all the analyses. Here, I switch between R and Python code chunks, but the package also allows objects to be used within both languages. Pretty cool.

To interact with Python via RMarkdowns in RStudio, check out the "reticulate" package website to get started: 

R Markdown Python Engine (aka the "reticulate" package):
https://rstudio.github.io/reticulate/articles/r_markdown.html

```{r reticulate, include=FALSE, message=F, warning=FALSE, results = "hide", echo=F}

# library to use Python within R:
library(reticulate)
# py_install("pandas")
#
```

3. Set up a Python environment for your work.  Here is why that is important:

Python environments:
"When installing Python packages it’s best practice to isolate them within a Python environment (a named Python installation that exists for a specific project or purpose). This provides a measure of isolation, so that updating a Python package for one project doesn’t impact other projects. The risk for package incompatibilities is significantly higher with Python packages than it is with R packages, because unlike CRAN, PyPI does not enforce, or even check, if the current versions of packages currently available are compatible."
https://rstudio.github.io/reticulate/articles/python_packages.html

(Anaconda and reticulate both help the user keep specific environments for each Python "project" that one may be working on.)

Here is another helpful link:

Primer on Python for R users:
https://rstudio.github.io/reticulate/articles/python_primer.html


```{r PythonEnvSetup}
# A Python virtual environment is a basic tool to isolate project dependencies within a Python installation, while a Conda environment is a more comprehensive package manager that not only creates isolated environments but also allows for managing complex dependencies across different programming languages

## Uncomment the code below and run once, the first time you work through the tutorial:

# # create a new environment
virtualenv_create("r-reticulate-BayesInPython")
# 
# # install SciPy
# virtualenv_install("r-reticulate", "scipy")
# 
# # import SciPy (it will be automatically discovered in "r-reticulate")
# scipy <- import("scipy")
# 
# # indicate that we want to use a specific virtualenv
# use_virtualenv("r-reticulate")
# 
# # import SciPy (will use "r-reticulate" as per call to use_virtualenv)
# scipy <- import("scipy")

## (Uncomment this code instead if you prefer a Conda environment) 
# # create a new environment 
# conda_create("r-reticulate")
# 
# # install SciPy
# conda_install("r-reticulate", "scipy")
# 
# # import SciPy (it will be automatically discovered in "r-reticulate")
# scipy <- import("scipy")
# 
# # indicate that we want to use a specific condaenv
# use_condaenv("r-reticulate")
# 
# # import SciPy (will use "r-reticulate" as per call to use_condaenv)
# scipy <- import("scipy")
# 

# reticulate::repl_python()

```


### Python modules

Python libraries or packages are called _modules_. To use them with reticulate, you first have to import them. The lines in the code chunk below must be called exactly once each for each new python module used; it does not need to re-run.

```{python, eval = F}
# You may have to install the packages if you have not used them before:
!pip install plotnine

```
```{r, eval = F}
# 
# # install packages for new environment
virtualenv_install("r-reticulate-BayesInPython", "pandas")
# 
# 
# # indicate that we want to use a specific virtualenv
use_virtualenv("r-reticulate-BayesInPython")
# 
# # import the modules (will use "r-reticulate" as per call to use_virtualenv)
pandas <- import("pandas")

```


## Set up

When completing an analysis, I may skip around in my script quite a bit. Therefore, I like all packages loaded, and all data uploaded and manipulated at the start of my script.

```{python readdatP}

import pandas as pd

# read in the moth data:
datP = pd.read_csv("data/moths.csv", delimiter = ",")
# drop first column, is just unnecessary indexing:
datP = datP.drop(datP.columns[0], axis=1)

```
```{python data_manip, results = "hide"}

# check data values: 
datP.info()

pd.set_option('display.max_columns', None) # Display all columns
datP.describe(percentiles=[0.5], include = 'all')

for col in datP:
  if (datP[col].dtype == 'object'):
    print(col)
    print(datP[col].unique())

# convert dates to dates in case further manipulation on them is needed:
cols = ['TransplantDate', 'DispInstallDate', 'TrapInstallDate', 'SamplingDate']
for col in cols:
    datP[col] = pd.to_datetime(datP[col])
datP['SamplingDate'].describe()

# standardize the counts
datP['mothsperday'] = datP['nYSB'] / datP['DaysOfCatch']

# datR <- datP %>%
#   mutate(TreatmentF = factor(Treatment), # if we need to fix levels in non-alpha order
         # LocationF = as.factor(Location),
         
# convert sampling date to string for nested RE's:
datP['SamplingDateC'] = str(datP['SamplingDate'])
# convert sampling date to numeric starting at 1 for GP's:
datP['numericDate'] = datP['SamplingDate'].dt.dayofyear
datP['numericDate'] = datP['numericDate'] - datP['numericDate'].min() + 1
datP['numericDate'].describe()

## average counts for each location, date
mean_cts = datP.groupby(['Location', 'Treatment', 'AssessmentNumber', 'SamplingDate', 'DATI'], as_index=False).agg({'nYSB': 'mean', 'mothsperday': 'mean'})
mean_cts.describe(include='all')

```
```{python functions}
# calculate trapping reduction for frequentist model fits.
def getTR(mod):
    """Calculate TR with confidence interval for stats model"""
    # Function body
    import pandas as pd

    b1 = mod.params.iloc[1]
    seb1 = mod.bse.iloc[1]
    
    outTR = {'TR': [round(100 * 1 - math.exp(b1))], 
    'lowerCI': [round(100 * 1 - math.exp(b1 + 1.96*seb1))],
    'upperCI': [round(100 * 1 - math.exp(b1 - 1.96*seb1))]}
    df = pd.DataFrame(outTR)
    return df

```

## Outline

* Introduce the data to be used throughout the tutorial.

* Fit generalized linear models (GLMs, here Poisson and negative binomial regressions). Models are fit in both frequentist and Bayesian frameworks to lay the groundwork for the more complicated models.

* Fit generalized linear models with random effects (GLMMs) to allow for correlations in space and time. Models are fit in both frequentist and Bayesian frameworks.

## Data Exploration

### Moth counts

Here is what the data look like. Columns: Location, Treatment, TrapID, Longitude, Latitude, AssessmentNumber, and SamplingDate should be self-explanatory... TransplantDate is the date that the rice seedlings were transplanted into the field; DispInstallDate is the treatment installation date; TrapInstallDate is the trap installation date; "nYSB" are the moth counts (number of YSB moths); DaysOfCatch is the number of days since the trap was previously sampled  or the number of days since trap installation if it is assessment number 1; DATI is days after trap installation; DADI is days after dispenser installation; and DAT is days after transplant. Some of these columns will be discussed more later.

```{python display_data}
# check data values: 
datP.info()

pd.set_option('display.max_columns', None) # Display all columns
datP.describe(percentiles=[0.5], include = 'all')

for col in datP:
  if (datP[col].dtype == 'object'):
    print(col)
    print(datP[col].unique())

```

The data are the moth counts collected at each trap (average per day), within each location and throughout the season. Note how the y-axis changes for each plot in the figure.

```{python plotMoths, fig.width = 8, fig.height = 6}
from plotnine import * # using ggplot in Python

mycolors = ["#9E0142", "#66C2A5"]
pmoths = (ggplot(datP, aes(x = 'DATI', y='mothsperday', color = 'Treatment')) +
        geom_jitter(height = 0, width = 0.75, alpha = 0.5) +
        geom_smooth(se=False) +
        facet_wrap(['Location'], scales = "free_y") +
        scale_color_manual(values = mycolors) +
        labs(title = "Male YSB moth counts for each location and trap",
             y = "Moth counts per trap per day",
             x = "Days after trap installation (DATI)",
             color = "Treatment"))
pmoths.draw(True)
```

### Trap locations

Trial locations in relation to each other. Some trial locations are closer to each other than others.

```{python}

# get API key (keep hidden though!)
import os
api_key = os.getenv("ggmapKey")

# import gmplot package
import gmplot

gmap1 = gmplot.GoogleMapPlotter(datP['Latitude'].mean(),
                                datP['Longitude'].mean(), 7, apikey=api_key )

gmap1.scatter(datP['Latitude'], datP['Longitude'], color='blue', size=2000, marker=False)

# Pass the absolute path
gmap1.draw( "test_map.html" )
#
```

Within each location, the treatment traps have a slightly different alignment:

```{python plotTrapLocs}

coords = datP[['Location', 'Treatment', 'TrapID', 'Latitude', 'Longitude']].drop_duplicates()
coords.describe(include = 'all')


pcoords = (ggplot(datP, aes(x = 'Longitude', y='Latitude', color = 'Treatment')) +
  geom_point() + 
  facet_wrap(['Location'], scales = "free") +
  scale_color_manual(values = mycolors) +
  theme(aspect_ratio=1,
        axis_text_x=element_blank(),
        axis_text_y=element_blank())  +
  geom_text(aes(label = 'TrapID'), nudge_x = 0.0002, nudge_y = 0.0002, color = "black", size = 8) +
  scale_x_continuous(expand = (0.3, 0)) +
  scale_y_continuous(expand = (0.3, 0)) +
  labs(title = "Trap ID's at each location"))
pcoords.draw(True)

```


### Timeline

```{python timeline, results="hide", fig.width = 7}

dates = datP.groupby(['Location', 'TransplantDate', 'TrapInstallDate'], as_index = False).agg(minDate=('SamplingDate', lambda x: x.min()), 
maxDate=('SamplingDate', lambda x: x.max()))
dates.describe(include = 'all')
dates = dates.sort_values(by=['minDate', 'Location'])
# dates$Location <- factor(dates$Location, levels = unique(dates$Location))

pdates = (ggplot(dates) +
  geom_errorbarh(aes(xmin = 'minDate', xmax = 'maxDate', 
                     y = 'Location', color = 'Location'),
                 height = 0.5, size = 1.2) +
  geom_jitter(aes('TransplantDate', y = 'Location'), width = 0, height = 0.1, size = 1.25) +
  geom_jitter(aes('TrapInstallDate', y = 'Location'), width = 0, height = 0.1, 
              size = 1.25, color = "gray") +
  labs(y = "Location",
       x = "Date range of moths data collection (2023)",
       title = "Trapping date ranges of each location",
       caption = "Black points are transplant dates; Grey points are trap installation dates") )
  
pdates.draw(True)

```

Note that not all traps are installed on exactly the same day; not all locations have rice transplanted on exactly the same day; and not all traps are sampled on exactly the same day.

## Basic GLM models

To start, we ignore all the spatial and temporal relationships in the data and assume each data point is independent and identically distributed (iid). **This is not a good model for the data!** It is just our starting off point.

When building hierarchical Bayesian models, it is always a good idea to start simple, make sure everything works as expected, and then build up.

### Mathematical model

Our response variable is a count, therefore we use either a Poisson or negative binomial distribution as the basis for our model. These models are known as generalized linear models (GLMs).
 
The model must also include an offset to account for the varying time intervals between sample. (If the traps are left unchecked for more days, we naturally expect the trap to have more moths. This varying effort is accounted for by the offset.)

#### Poisson regression model

We start with a poisson version of our regression model:

$$
y_i \sim Pois(\lambda_i) \\
log(\lambda_i ) = \beta_0 + \beta_1 x_{A,i} + \mbox{ln} \left(DaysOfCatch_i\right)
$$

where, for data samples $i = 1, ..., N$, 

$y_i$ is moth count $i$,

$\lambda_i$ is the expected moth count for sample $i$,

$\beta_0$ is the intercept of the model. Here, it is the basis for the expected moth count for our control treatment.

$\beta_1$ is the expected difference in moth counts between the control group and the treatment A group, 

$x_{i}$ is an indicator variable that equals 1 if sample $i$ is from a treatment A field and equals 0 otherwise, and

$\left(DaysOfCatch_i\right)$ is the offset to account for the varying time interval between samples. This is necessary because with a longer time interval, more moths will fly into the trap.

#### Negative binomial regression model

In ecology, there is always additional variability in our data than what the Poisson distribution allows. To account for the additional variability in our model, we switch to a negative binomial distribution:

$$
y_i \sim NegBinom(\lambda_i, \phi) \\
log(\lambda_i ) = \beta_0 + \beta_1 x_{i} + \mbox{ln} \left(DaysOfCatch_i\right)
$$

"Negative binomial regression is used to model count data for which the variance is higher than the mean." (https://www.pymc.io/projects/examples/en/latest/generalized_linear_models/GLM-negative-binomial-regression.html)

For the negative binomial regression, all variables and parameters are defined as before, but the model has an additional parameter, $\phi$, that allows for the additional variation. The negative binomial distribution and regression model can be written in many ways. In our models, $\phi$ is defined through the following formula where the variance associated with an expected moth count, $\lambda$, is:

$$
Var(\lambda) = \lambda + \phi \cdot\lambda^2
$$

(For a Poisson distribution, $Var(\lambda) = \lambda$.)

#### Bayesian mathematical model

I also fit these models using a Bayesian framework, again to build our foundation for the more complex models that come later.

For the Bayesian models, I only show the more complex negative binomial version.

To make our models Bayesian, we add priors to our parameters:

$$
y_i \sim NegBinom(\lambda_i, \phi) \\
log(\lambda_i ) = \beta_0 + \beta_1 x_{A,i} + \mbox{ln} \left(DaysOfCatch_i\right) \\
\beta_0 \sim Normal(0, 2) \\
\beta_1 \sim Normal(0, 2) \\
\phi \sim HalfCauchy(0,1)
$$

The models use slightly informative priors. When working on a log-scale (e.g., with Poisson or negative binomial distributions), this often becomes essential to avoid parameter estimates on the boundaries. (See Hooten and Hobbs for a good discussion on this issue.)

#### Trapping reduction defined

The primary metric of interest from these models is a derived parameter that we call trapping reduction (TR):

$$
TR = 100 - 100 e ^{(-\beta_1)}
$$

To obtain a 95% confidence interval for TR (for the frequentist version of our model), we use the following approximation:

$$
TR = 100 - 100 e ^{(-\beta_1 \pm 2 \cdot SE(\beta_1))}
$$

For the Bayesian version, the derived parameter is sampled as part of the MCMC iterations, and the 95% credible interval is the 2.5%, 97.5% quantiles of its resulting distribution.

### Fitting the frequentist models

Note: when I predict for new data, I set DaysOfCatch = 1, and then I compare to the moths per day variable (mothsperday = nYSB / DaysOfCatch). I want to exclude any patterns related to the varying time intervals.

These models show a very tight confidence intervals for our predictions. But also, the residual deviance is MUCH greater than the degree of freedom, indicating a lack of fit. The plot of the predictions overlaid on the data also demonstrate the lack of fit.

These models are fit using the 'statsmodels' module because of its parallel syntax to R.

```{python pois1py}

import numpy as np
from patsy import dmatrices # for dmatrices function
import statsmodels.api as sm # for regressions
import math

# POISSON regression
form = """ nYSB ~ Treatment"""
# X = moths["Treatment"]
# y = moths["nYSB"]
y, X = dmatrices(form, datP, return_type = 'dataframe')
mod1p = sm.GLM(y, X, 
              offset=np.log(datP["DaysOfCatch"]),
              family = sm.families.Poisson())
results1p = mod1p.fit()
print(results1p.summary())

TR_1p = getTR(results1p)
print(TR_1p)

# predict from poisson regression:
preds1p = results1p.get_prediction(X, which = 'linear')
preds1pDF = preds1p.summary_frame() #summary_frame() returns a pandas DF
preds1pDF.describe()

preds1pDF = datP
preds1pDF['exp_values'] = preds1pDFtmp['predicted'].apply(math.exp)
preds1pDF['lowerCI'] = preds1pDFtmp['ci_lower'].apply(math.exp)
preds1pDF['upperCI'] = preds1pDFtmp['ci_upper'].apply(math.exp)

## plot predicted and actual:
plot1p = (ggplot(preds1pDF, 
                aes(x = 'DATI', y='exp_values', color = 'Treatment')) +
        geom_line() +
        geom_ribbon(aes(ymin = 'lowerCI', ymax = 'upperCI', fill = 'Treatment'),
              alpha = 0.3) +
        geom_point(aes(x='DATI', y='mothsperday', color = 'Treatment'), 
            data = mean_cts, size = 1.1) +
        facet_wrap(['Location'], scales = "free_y") +
        scale_color_manual(values = mycolors) +
        scale_fill_manual(values = mycolors) +
        labs(title = "Poisson GLM predictions",
             subtitle = "Lines are predictions, points are the data"))
plot1p.draw(True)

```


The model is such a bad fit to the data, it is hard to even tell what is going on in the plot above.

For the negative binomial model, our confidence intervals are a little wider, but we are still ignoring all the correlations in our data. And the plot of the predictions again indicates the lack of fit.

```{python nb1py}

from scipy.stats import nbinom

form = """ nYSB ~ Treatment"""
y, X = dmatrices(form, datP, return_type = 'dataframe')
mod1nb = sm.NegativeBinomial(y,X, offset=np.log(datP["DaysOfCatch"]))
results1nb = mod1nb.fit()
print(results1nb.summary())

TR_1nb = getTR(results1nb)
print(TR_1nb)

# predict from NB regression:
preds1nb = results1nb.get_prediction(X, which = 'linear')
preds1nbDFtmp = preds1nb.summary_frame()
preds1nbDFtmp.describe()

preds1nbDF = datP
preds1nbDF['exp_values'] = preds1nbDFtmp['predicted'].apply(math.exp)
preds1nbDF['lowerCI'] = preds1nbDFtmp['ci_lower'].apply(math.exp)
preds1nbDF['upperCI'] = preds1nbDFtmp['ci_upper'].apply(math.exp)




## plot predicted and actual:
plot1nb = (ggplot(preds1nbDF, aes(x = 'DATI', y='exp_values', color = 'Treatment')) +
        geom_line() +
        geom_ribbon(aes(ymin = 'lowerCI', ymax = 'upperCI', fill = 'Treatment'),
              alpha = 0.3) +
        geom_point(aes(x='DATI', y='mothsperday', color = 'Treatment'), 
            data = mean_cts, size = 1.1) +
        facet_wrap(['Location'], scales = "free_y") +
        scale_color_manual(values = mycolors) +
        scale_fill_manual(values = mycolors) +
        labs(title = "Negative binomial GLM predictions",
             subtitle = "Lines are predictions, points are the data"))

plot1nb.draw(True)


```

### Bayesian (bambi) models-- Python

"Bambi" is the Python module that is *very* similar to "brms" package in R. I'm pretty sure the defaults use different algorithms though (which may affect your results), so be sure to read up on the details of both!

Because the models take a few minutes to fit, I usually fit them when I am initially running through my code, save them, and then only load the model fit when rendering the Rmarkdown file and creating the resulting figures.

The prior_summary command is helpful if you do not know what a parameter is called in 'brms'. Here, I used the command to find out what they called their $\phi$ parameter. (They call it the shape parameter.)

```{python nbBambi1}
# https://bambinos.github.io/bambi/notebooks/negative_binomial.html
import bambi as bmb
bambi1nb = bmb.Model("nYSB ~ Treatment + offset(log(DaysOfCatch))", 
                    datP, family="negativebinomial",
                    priors = {
                        "Intercept": bmb.Prior("Normal", mu = 0, sigma = 2),
                        "Treatment": bmb.Prior("Normal", mu = 0, sigma = 2),
                        "alpha": bmb.Prior("HalfCauchy", beta = 1)
                    })
resbambi1nb = bambi1nb.fit(tune = 500, draws=2000, chains=6, random_seed = 5000)
# tune = warmup; draws = iter

import pickle # to save bambi object to file.
filename = 'output/resbambi1nb.pkl'
# save bambi object to file.
with open(filename, 'wb') as file:
    pickle.dump(resbambi1nb, file)
 
```

The Bayesian parameter estimates match very closely to the frequentist estimates. This is expected, but reassuring that we have built the correct foundation for the more complicated models to come. The first summary output is from the Bayesian model; the second is from the frequentist fit. (Note that in the fequentitst fit, they summarize the inverse shape parameter compared to the bambi fit.)

Comparisons of TR predictions and predicted moth count values can be found at the end of the page.

```{python nbBambiOutput1}
import arviz as az
import matplotlib.pyplot as plt 
import pickle

# Load the object from the file
filename = 'output/resbambi1nb.pkl'
with open(filename, 'rb') as file:
    resbambi1nb = pickle.load(file)

# to see the priors:
# bambi1nb

# summary stats
print(az.summary(resbambi1nb))
print(results1nb.summary())


# CALC TR!
b1 = resbambi1nb.posterior["Treatment"].stack(sample=("chain", "draw")).to_dataframe(name = "trtA")
import numpy as np

TR1nb= 100 * (1 - np.exp(b1['trtA']))
TR1nb.quantile(q=[0.025, 0.5, 0.975])
outTR1nb = {'TR': [round(TR1nb.quantile(q=0.5))], 
    'lowerCI': [round(TR1nb.quantile(q=0.025))], 
    'upperCI': [round(TR1nb.quantile(q=0.975))]}

# post = idata.posterior
# post
# post["log_tau"] = np.log(post["tau"])
# idata.posterior
# post.mean() # to get mean TR
# post["mlogtau"] = post["log_tau"].rolling({"draw": 50}).mean()


# plot predictions!!!
# predictions
# fig, ax = plt.subplots(figsize=(7, 3), dpi=120)
# bmb.interpret.plot_predictions(bambi1nb, resbambi1nb, average_by = ["Treatment"]); #,"Village"]);

newdata = datP
newdata['DaysOfCatch'] = 1

# Plot the data
idata = bambi1nb.predict(resbambi1nb, kind="response_params", data=newdata, inplace = False)
# idata = model.predict(fitted, kind='pps', inplace=False)


# stacked = az.extract(idata)
# stacked
# stacked.mu.values
# idata.observed_data.nYSB

# bayes_p1nb <- ggplot(tmp, 
#        aes(x = DATI, y = Estimate, color = TreatmentF)) + 
#   geom_line() + 
#   geom_ribbon(aes(ymin = `Q2.5`, ymax = `Q97.5`, fill = TreatmentF),  # regression line and CI
#                     alpha = 0.3) +
#   geom_point(data = mean_cts, 
#              aes(DATI, mean_mothsperday, color = TreatmentF), size = 1.1) +
#   facet_wrap(~Location, scales = "free_y") + 
#   scale_color_manual(values = mycolors[]) +
#   scale_fill_manual(values = mycolors[]) +
#   labs(title = "Bayesian neg. binomial GLM predictions",
#        subtitle = "Lines are predictions, points are the data")
# print(bayes_p1nb)


```

### Bayesian (PyMC) models-- Python

For most needs, the "bambi" module should suffice. However, when I want to build my complicated non-linear function with multi-dimensional Gaussian Processes, I am going to need a more foundational module-- PyMC. Therefore, I will also build all models using this module. Learning PyMC will also help with extracting derived parameters and predictions, etc. from bambi

```{python nbPyMC1, eval = F}
# NB Bayes:
#https://discourse.pymc.io/t/biased-results-from-poisson-regression-model/4349/5
import numpy as np
import pandas as pd

from formulae import design_matrices
import pymc as pm
# Create the design matrix using formulae
# Wrapping the variable in 'C()' enforces categorical values, with the first entry as the reference
dm = design_matrices('nYSB ~ C(Treatment)', data=datP, na_action='error')
# print(dm.common.as_dataframe().head())

# Expand out into X and y
X = dm.common.as_dataframe()
y = dm.response.as_dataframe().squeeze() # This ensures y is 1-dimensional. 

DaysOfCatch = datP['DaysOfCatch']

# Create labelled coords which help the model keep track of everything(i.e., use factor names)
c = {'predictors': X.columns, # Model now aware of number of and labels of predictors
     'N': range(len(y))} # and number of observations

with pm.Model(coords = c) as model_pymc1nb:
    # Set the data, and convert to NumPy arrays (to use matrix multiplication
    Xdata = pm.Data('Xdata', X.to_numpy())
    Ydata = pm.Data('Ydata', y.to_numpy())
    
    # priors
    # factor_coefs = pm.Normal('factor_coefs', mu=0, sigma=2, shape=len(X.columns))    
    beta = pm.Normal('beta', mu=0, sigma=2, dims = 'predictors')
    alpha = pm.HalfCauchy('alpha', beta = 1)

    # TR derived parameter:
    TR = pm.Deterministic('TR', 100 - 100 * np.exp(beta[1]))
    
    # likelihood:
    mu = pm.math.exp(Xdata @ beta + # pm.math.dot(X, factor_coefs) + #@ operator is for matrix multiplication 
                     np.log(DaysOfCatch)) 
    pm.NegativeBinomial('y', mu=mu, alpha = alpha, observed=Ydata)
    
    # Sample according to your specifications, but tune for longer and take fewer draws
    pymc1nb = pm.sample(draws=1000, tune=3000, cores=4, random_seed=45, step=pm.NUTS())#Metropolis())
    # pymc1nb is called 'trace' in lots of the help files


import pickle # to save PyMC object to file.
filename = 'output/respymc1nb.pkl'
# save bambi object to file.
with open(filename, 'wb') as file:
    pickle.dump(pymc1nb, file)


```

```{python nbPyMCOutput1}

# Load the object from the file
import pickle
filename = 'output/respymc1nb.pkl'
with open(filename, 'rb') as file:
    pymc1nb = pickle.load(file)

# summary stats
print(az.summary(pymc1nb))

# CALC TR!
outTR1nb = round(TR1nb.quantile(q=[0.025, 0.5, 0.975]).to_pandas())

```


## GLMM: Random effect (RE) for locations

The first fix we make to the model is acknowledging that overall average  moth pressure varies from location to location (see Fig 1 of moth counts). To make this fix, we add a location random effect (RE) and our model becomes a generalized linear mixed-effects model (GLMM or GLMER).

We also want to acknowledge that the treatment effect may vary from location to location-- sometimes we see a big difference in moth counts between control and treatment fields, and sometimes the difference is smaller. For inference though, we are only interested in the larger picture, which is the overall trapping reduction. (We are not interested in what happens at these exact locations per se, we are more interested in the average treatment effect.) Therefore, we also add a treatment random effect.

Because dates will have different affect at different location (partially due to different planting dates), they are nested and not crossed.

(Should date be planting date and not sampling date? )

### GLMM mathematical model

I only show the Poisson version of the model, the NB version is a straightforward extension.

$$
y_{ik} \sim Pois(\lambda_{ik}) \\
log(\lambda_{ik} ) = \beta_0 + \beta_1 x_{A,i} + \gamma_{0k} x_{ik}  + \gamma_{1k} x_{i} x_{ik} + \mbox{ln} \left(DaysOfCatch_i\right) \\
\gamma_{0k} \sim Normal(0, \sigma^2_{0}) \\
\gamma_{1k} \sim Normal(0, \sigma^2_{1}) \\
$$

where, in addition to the variables and parameters defined for the GLM, we have

$k = 1,..., K=10$ represents the locations, 

$y_{ik}$ is moth count $i$ from location $k$,

$\lambda_{ik}$ is the expected moth count for sample $i$  from location $k$,

$\beta_0$ is the expected moth count for our control treatment for a new location,

$\beta_1$ is the expected difference in moth counts between the control group and the treatment A group for a new location, 

$\gamma_{0k}$ is the random intercept associated with location $k$, which leads to different background moth pressures at each location. All $\gamma_{0k}$ come from an iid Normal distribution.

$\gamma_{1k}$ is the random slope associated with location $k$, which leads to different treatment effects at each location. All $\gamma_{1k}$ come from an iid Normal distribution, and

$x_{ik}$ is an indicator variable that equals 1 if moth count $i$ is associated with location $k$ and 0 otherwise.

OK, technically I should be introducing matrices here because $k = 1,..., K=10$ locations, so we need a new indicator variable for each location, but I'm going to be lazy and write it like this for now.


#### Bayesian GLMM

Here is the same model with a Bayesian framework and using a negative binomial regression:

To make our models Bayesian, we add priors to our parameters:

$$
y_{ik}\sim NegBinom(\lambda_{ik}, \phi) \\
log(\lambda_{ik} ) = \beta_0 + \beta_1 x_{A,i} + \gamma_{0k} x_{ik}  + \gamma_{1k} x_{A,i} x_{ik} + \mbox{ln} \left(DaysOfCatch_i\right) \\ 
\gamma_{0k} \sim Normal(0, \sigma^2_{0}) \\
\gamma_{1k} \sim Normal(0, \sigma^2_{1}) \\
\beta_0 \sim Normal(0, 2) \\
\beta_1 \sim Normal(0, 2) \\
\sigma^2_{0} \sim HalfCauchy(0,1) \\
\sigma^2_{1} \sim HalfCauchy(0,1) \\
\phi \sim HalfCauchy(0,1) \\
$$


#### Bayesian GLMM -- Matrix version

I DO write the Bayesian version in matrix form because it matches the coding (and is more correct given the need for a $KxK$ $Z$ design matrix):

$$
\mathbf{Y} \sim NegBinom(\boldsymbol\lambda, \phi) \\
log(\boldsymbol\lambda) = \mathbf{X}\boldsymbol\beta + \mathbf{Z}\boldsymbol\gamma  + \mbox{ln} \left(\bf{DaysOfCatch}\right) \\
\boldsymbol\beta \sim Normal(\mathbf{0}, 2\mathbf{I}) \\
\boldsymbol\gamma \sim Normal(\mathbf{0}, \boldsymbol\sigma^2 \mathbf{I}) \\
\boldsymbol\sigma^2 \sim HalfCauchy(\mathbf{0}, 1\mathbf{I})) \\
\phi \sim HalfCauchy(0,1)
$$


### GLMM frequentist fit

Python is a little disappointing for GLMMs. We use "pymer4" module which is just a wrapper to the R library "lme4". This is awkward here-- we are wrapping Python code inside R, and the Python code is a wrapper back to R. So, I get a warning. I am including the code here anyways for anyone working outside of reticulate and RStudio.

(If we only have random intercepts, you can use the "statsmodels" module, but it cannot accommodate random intercepts and slopes, which is what we have.)

You can also use the "GPBoost" module, but it is overly complicated for our small data set.

The "pymer4" module installation is giving me some trouble, so I will fill in this section later.

<!-- When we plot our predictions, we see that we now have better estimates for the overall mean at each location, and we see how much they vary from location to location, but there is a strong temporal pattern at each location that we are missing. -->

```{python poisPy, eval = F}

from pymer4 import Lmer
from plotnine import *
modp2 = Lmer('nYSB ~ DetailsF + (DetailsF|Location) + offset(log(DaysOfCatch))', moddat, family = "poisson")
modp2.fit()
print(modp2.summary())

# predict from poisson regression:
# poisson_predictions = modp2.predict(moths, skip_data_checks=True, verify_predictions=False)

predsDF = moddat
predsDF['exp_values'] = modp2.data['fits'] # poisson_predictions # 

## plot predicted and actual:
pnbRE = (ggplot(predsDF, aes(x = 'DATI', y='exp_values', color = 'DetailsF')) +
        geom_point(aes(x = 'DATI', y = 'nYSB', color = 'DetailsF')) +
        geom_smooth(se=False) +
        facet_wrap(['Location'], scales = "free_y") +
        scale_color_manual(values = ["#9E0142", "#66C2A5"]) +
        labs(title = "Male YSB moth counts for each location and trap",
             y = "Moth counts per trap per day",
             x = "DAI",
             color = "Treatment"))

pnbRE.draw(True)

# Actual versus predicted:
# sns.regplot(x="fits", y="nYSB", data=modp2.data, fit_reg=True)
# plt.show()

# another way to view output:
# modp2.plot_summary()
# plt.show()

```

Google's Gemini AI is convinced that a negative binomial regression with random slopes and intercepts is possible in Python, but:

* the Lmer function does not allow for neg binomial regression models. 

* "statsmodels" similarly only supports binomial and Poisson models using Bayesian methods. 

Therefore, we only fit the negative binomial regression model within the Bayesian framework. This is a category where R takes the win over Python!


### GLMM Bayesian (bambi) models-- Python

By default, the prior distributions of random slopes and intercepts are correlated in "brm". If the correlation is non-significant, you may want to remove that correlation to simplify your model. To remove the correlation, change the random effects term from "(1 + TreatmentF|Location)" to " (1 + TreatmentF||Location)" (has an extra vertical line), which makes them independent. (TRUE FOR BAMBI??)

You'll notice that the parameter estimates are slightly different from the frequentist NB model fit here-- the more complicated your model is, the more likely you will find this is true with slightly different versions of your model (here, adding priors and fitting with a different algorithm). 

Again, notice the subtle differences between "brms" and "bambi".

```{python nbBambi2}
# https://bambinos.github.io/bambi/notebooks/negative_binomial.html
import bambi as bmb
bambi2nb = bmb.Model("nYSB ~ Treatment + (Treatment|Location) + offset(log(DaysOfCatch))", 
                    datP, family="negativebinomial",
                    priors = {
                        "Intercept": bmb.Prior("Normal", mu = 0, sigma = 2),
                        "Treatment": bmb.Prior("Normal", mu = 0, sigma = 2),
                        "alpha": bmb.Prior("HalfCauchy", beta = 1),
                        "1|Location": bmb.Prior("Normal", mu = 0, sigma = bmb.Prior("HalfCauchy", beta = 1)),
                        "Treatment|Location": bmb.Prior("Normal", mu = 0, sigma = bmb.Prior("HalfCauchy", beta = 1))
                    })
resbambi2nb = bambi2nb.fit(target_accept = 0.90, tune = 500, draws=2000, chains=6, random_seed = 5000)
# 

import pickle # to save bambi object to file.
filename = 'output/resbambi2nb.pkl'
# save bambi object to file.
with open(filename, 'wb') as file:
    pickle.dump(resbambi2nb, file)
    
    
```
```{python nbBambiOutput2}
import arviz as az
import matplotlib.pyplot as plt 
import pickle

# Load the object from the file
filename = 'output/resbambi2nb.pkl'
with open(filename, 'rb') as file:
    resbambi2nb = pickle.load(file)
    

# to see priors:
bambi2nb

# summary stats
print(az.summary(resbambi2nb,  # kind="stats"))
                 var_names=["Intercept", "Treatment", "alpha", "1|Location_sigma", "Treatment|Location_sigma"]))

# quick diagnostics
az.plot_trace(resbambi2nb, backend_kwargs={"layout": "constrained"}, legend=True,
var_names = ["Intercept", "Treatment", "alpha", "1|Location_sigma", "Treatment|Location_sigma"]);
plt.show()

az.plot_trace(resbambi2nb, backend_kwargs={"layout": "constrained"}, legend=True, 
var_names = ["1|Location", "Treatment|Location"]);
plt.show()

```
```{python nbBambiOutput2b}

# CALC TR!
b1 = resbambi2nb.posterior["Treatment"].stack(sample=("chain", "draw")).to_dataframe(name = "trtA")
import numpy as np

TR2nb= 100 * (1 - np.exp(b1['trtA']))
TR2nb.quantile(q=[0.025, 0.5, 0.975])
outTR2nb = {'TR': [round(TR2nb.quantile(q=0.5))], 
    'lowerCI': [round(TR2nb.quantile(q=0.025))], 
    'upperCI': [round(TR2nb.quantile(q=0.975))]}
print(outTR2nb)

# post = idata.posterior
# post
# post["log_tau"] = np.log(post["tau"])
# idata.posterior
# post.mean() # to get mean TR
# post["mlogtau"] = post["log_tau"].rolling({"draw": 50}).mean()


# plot predictions!!!
# predictions
# fig, ax = plt.subplots(figsize=(7, 3), dpi=120)
# bmb.interpret.plot_predictions(bambi1nb, resbambi1nb, average_by = ["Treatment"]); #,"Village"]);

newdata = datP
newdata['DaysOfCatch'] = 1

# Plot the data
idata = bambi2nb.predict(resbambi2nb, kind="response_params", data=newdata, inplace = False)
# idata = model.predict(fitted, kind='pps', inplace=False)


# stacked = az.extract(idata)
# stacked
# stacked.mu.values
# idata.observed_data.nYSB

# bayes_p2nb <- ggplot(tmp, 
#        aes(x = DATI, y = Estimate, color = TreatmentF)) + 
#   geom_line() + 
#   geom_ribbon(aes(ymin = `Q2.5`, ymax = `Q97.5`, fill = TreatmentF),  # regression line and CI
#                     alpha = 0.3) +
#   geom_point(data = mean_cts, 
#              aes(DATI, mean_mothsperday, color = TreatmentF), size = 1.1) +
#   facet_wrap(~Location, scales = "free_y") + 
#   scale_color_manual(values = mycolors[]) +
#   scale_fill_manual(values = mycolors[]) +
#   labs(title = "Bayesian neg. binomial GLM predictions",
#        subtitle = "Lines are predictions, points are the data")
# print(bayes_p2nb)


```


### GLMM Bayesian (PyMC) models-- Python

Notice that my model framework is different from the tutorial within PyMC. In PyM tutorial, our beta's are the means of the random intercept and slope. Here, I explicitly keep them as separate parameters. Is their way faster than mine? Perhaps that is why they do it that way? I'm sticking with this way because this is how I have always built the more complicated models that are forthcoming. Also, it is helpful to see similar models coded in different ways-- helps you become more familiar with the programming and betters your understanding of the model itself.

```{python nbPyMCsetup2}
# NB Bayes:
#https://discourse.pymc.io/t/biased-results-from-poisson-regression-model/4349/5
import numpy as np
import pandas as pd
import patsy # for dmatrix function

from formulae import design_matrices
import pymc as pm
# Create the design matrix using formulae
# Wrapping the variable in 'C()' enforces categorical values, with the first entry as the reference
dm = design_matrices('nYSB ~ C(Treatment)', data=datP, na_action='error')
# print(dm.common.as_dataframe().head())

# Expand out into X and y
X = dm.common.as_dataframe()
y = dm.response.as_dataframe().squeeze() # This ensures y is 1-dimensional. 

# https://github.com/JackCaster/GLM_with_PyMC3/blob/master/notebooks/LME%20-%20Sleep%20study.ipynb
Z = patsy.dmatrix('1 + Location', data=datP, return_type='dataframe')
Z.intercept = 0 # to match our model!
# print(Z)

n_obs = datP.shape[0]
nlocs = Z.shape[1]

DaysOfCatch = datP['DaysOfCatch']

# Create labelled coords which help the model keep track of everything(i.e., use factor names)
c = {'predictors': X.columns, 
     # 'nlocs': Z.shape[1],
     # 'n_obs': moddat.shape[0],
     'N': range(len(y))}

```
```{python nbPyMCrun2, eval = F}

with pm.Model(coords = c) as model_pymc2nb:
    # priors
    sigma0 = pm.HalfCauchy('sigma0', beta = 1)
    gamma0 = pm.Normal('gamma0', mu=0, sigma = sigma0, shape = len(Z.columns))
    sigma1 = pm.HalfCauchy('sigma1', beta = 1)
    gamma1 = pm.Normal('gamma1', mu=0, sigma = sigma1, shape = len(Z.columns))
    beta_coefs = pm.Normal('beta_coefs', mu=0, sigma=2, shape=len(X.columns))
   
    alpha = pm.HalfCauchy('alpha', beta = 1)

    # TR derived parameter:
    TR = pm.Deterministic('TR', 100 - 100 * np.exp(beta_coefs[1]))

    # likelihood:
    mu = pm.Deterministic(
        'mu',
        pm.math.dot(X, beta_coefs) + 
        pm.math.dot(Z, gamma0) +
        pm.math.dot(Z, gamma1) +
        np.log(DaysOfCatch)
    )

    # This set-up would be an over-dispersed poisson:
    # log_rate = pm.Normal('log_rate', mu=mu, sigma=alpha, shape=n_obs)
    # predy = pm.Poisson('predy', mu=pm.math.exp(log_rate), observed=y)

    # This is the neg binomial set-up:
    pm.NegativeBinomial('predy', mu=pm.math.exp(mu), alpha = alpha, observed=y)
    
    idata_pymc2nb = pm.sample(nuts={'target_accept':0.9}, draws=1000, tune=3000, cores=4, random_seed=45)

import pickle # to save PyMC object to file.
filename = 'output/idata_pymc2nb.pkl'
# save bambi object to file.
with open(filename, 'wb') as file:
    pickle.dump(idata_pymc2nb, file)


```
```{python nbPyMC1, eval = F}
# code kept here for comparison; to make above more efficient (eventually)
with pm.Model(coords = c) as model_pymc1nb:
    # Set the data, and convert to NumPy arrays (to use matrix multiplication
    Xdata = pm.Data('Xdata', X.to_numpy())
    Ydata = pm.Data('Ydata', y.to_numpy())
    
    # priors
    # factor_coefs = pm.Normal('factor_coefs', mu=0, sigma=2, shape=len(X.columns))    
    beta = pm.Normal('beta', mu=0, sigma=2, dims = 'predictors')
    alpha = pm.HalfCauchy('alpha', beta = 1)

    # TR derived parameter:
    TR = pm.Deterministic('TR', 100 - 100 * np.exp(beta[1]))
    
    # likelihood:
    mu = pm.math.exp(Xdata @ beta + # pm.math.dot(X, factor_coefs) + #@ operator is for matrix multiplication 
                     np.log(DaysOfCatch)) 
    pm.NegativeBinomial('y', mu=mu, alpha = alpha, observed=Ydata)
    
    # Sample according to your specifications, but tune for longer and take fewer draws
    pymc1nb = pm.sample(draws=1000, tune=3000, cores=4, random_seed=45, step=pm.NUTS())#Metropolis())
    # pymc1nb is called 'trace' in lots of the help files

```

Starting with these models, I specify which variables for which I want to see the posteriors and summary statistics. In general, we don't usually want the output for any latent variables as that would overwhelm our screens and are not the parameters of most interest right now.

```{python nbPyMCOutput2}

# Load the object from the file
import pickle
filename = 'output/idata_pymc2nb.pkl'
with open(filename, 'rb') as file:
    idata_pymc2nb = pickle.load(file)

# summary stats
print(az.summary(idata_pymc2nb, var_names=["beta_coefs", "alpha", "sigma0", "sigma1"])) 

# quick diagnostics
# az.plot_trace(idata_pymc2nb, backend_kwargs={"layout": "constrained"}, legend=True, var_names=["beta_coefs", "alpha", "sigma0", "sigma1"])
# plt.show()

# az.plot_trace(idata_pymc2nb, backend_kwargs={"layout": "constrained"}, legend=True, var_names=["gamma0", "gamma1"])
# plt.show() # many are bi-model--- not a good fit!!!

# CALC TR!
print(az.summary(idata_pymc2nb, var_names=["TR"])) 

```


## GLMM: RE for location, crossed SamplingDate

### Crossed GLMM mathematical model

Model to be filled in.

#### Bayesian crossed GLMM

Model to be filled in.

### Crossed GLMM frequentist fit

Code to be filled in.

### Crossed GLMM Bayes (bambi) version

```{python nbBambi3}
import bambi as bmb
bambi3nb = bmb.Model("nYSB ~ Treatment + +(Treatment|Location) +  (Treatment|Location:SamplingDateC) + offset(log(DaysOfCatch))", 
                    datP, family="negativebinomial",
                    priors = {
                        "Intercept": bmb.Prior("Normal", mu = 0, sigma = 2),
                        "Treatment": bmb.Prior("Normal", mu = 0, sigma = 2),
                        "alpha": bmb.Prior("HalfCauchy", beta = 1),
                        "1|Location:SamplingDateC": bmb.Prior("Normal", mu = 0, sigma = bmb.Prior("HalfCauchy", beta = 1)),
                        "Treatment|Location:SamplingDateC": bmb.Prior("Normal", mu = 0, sigma = bmb.Prior("HalfCauchy", beta = 1))
                    })
idata_bambi3nb = bambi3nb.fit(target_accept = 0.90, tune = 500, draws=2000, chains=6, random_seed = 5000)
# 

import pickle # to save bambi object to file.
filename = 'output/idata_bambi3nb.pkl'
# save bambi object to file.
with open(filename, 'wb') as file:
    pickle.dump(idata_bambi3nb, file)
    
    
```
```{python nbBambiOutput3}
import arviz as az
import matplotlib.pyplot as plt 
import pickle

# Load the object from the file
filename = 'output/idata_bambi3nb.pkl'
with open(filename, 'rb') as file:
    idata_bambi3nb = pickle.load(file)
    

# to see priors:
bambi3nb

## to see data variables:
idata_bambi3nb.posterior.data_vars

# summary stats
print(az.summary(idata_bambi3nb,  # kind="stats"))
                 var_names=["Intercept", "Treatment", "alpha", "1|Location:SamplingDateC_sigma", "Treatment|Location:SamplingDateC_sigma"]))

# quick diagnostics
az.plot_trace(idata_bambi3nb, backend_kwargs={"layout": "constrained"}, legend=True,
var_names = ["Intercept", "Treatment", "alpha", "1|Location:SamplingDateC_sigma", "Treatment|Location:SamplingDateC_sigma"]);
plt.show()

az.plot_trace(idata_bambi3nb, backend_kwargs={"layout": "constrained"}, legend=False, 
var_names = ["1|Location:SamplingDateC", "Treatment|Location:SamplingDateC"]);
plt.show()

```
```{python nbBambiOutput3b}

# CALC TR!
b1 = idata_bambi3nb.posterior["Treatment"].stack(sample=("chain", "draw")).to_dataframe(name = "trtA")
import numpy as np

TR3nb= 100 * (1 - np.exp(b1['trtA']))
TR3nb.quantile(q=[0.025, 0.5, 0.975])
outTR3nb = {'TR': [round(TR3nb.quantile(q=0.5))], 
    'lowerCI': [round(TR3nb.quantile(q=0.025))], 
    'upperCI': [round(TR3nb.quantile(q=0.975))]}
print(outTR3nb)


# plot predictions!!!

```

### Crossed GLMM Bayes (PyMC) version

```{python nbPyMCsetup3}
# NB Bayes:
#https://discourse.pymc.io/t/biased-results-from-poisson-regression-model/4349/5
import numpy as np
import pandas as pd
import patsy # for dmatrix function

from formulae import design_matrices
import pymc as pm
# Create the design matrix using formulae
# Wrapping the variable in 'C()' enforces categorical values, with the first entry as the reference
dm = design_matrices('nYSB ~ C(Treatment)', data=datP, na_action='error')
# print(dm.common.as_dataframe().head())

# Expand out into X and y
X = dm.common.as_dataframe()
y = dm.response.as_dataframe().squeeze() # This ensures y is 1-dimensional. 

# https://github.com/JackCaster/GLM_with_PyMC3/blob/master/notebooks/LME%20-%20Sleep%20study.ipynb
Z = patsy.dmatrix('1 + Location', data=datP, return_type='dataframe')
Z.intercept = 0 # to match our model!
# print(Z)

n_obs = datP.shape[0]
nlocs = Z.shape[1]

DaysOfCatch = datP['DaysOfCatch']

# Create labelled coords which help the model keep track of everything(i.e., use factor names)
c = {'predictors': X.columns, 
     # 'nlocs': Z.shape[1],
     # 'n_obs': moddat.shape[0],
     'N': range(len(y))}

```
```{python nbPyMCrun3, eval = F}

with pm.Model(coords = c) as model_pymc3nb:
    # priors
    sigma0 = pm.HalfCauchy('sigma0', beta = 1)
    gamma0 = pm.Normal('gamma0', mu=0, sigma = sigma0, shape = len(Z.columns))
    sigma1 = pm.HalfCauchy('sigma1', beta = 1)
    gamma1 = pm.Normal('gamma1', mu=0, sigma = sigma1, shape = len(Z.columns))
    beta_coefs = pm.Normal('beta_coefs', mu=0, sigma=2, shape=len(X.columns))
   
    alpha = pm.HalfCauchy('alpha', beta = 1)

    # TR derived parameter:
    TR = pm.Deterministic('TR', 100 - 100 * np.exp(beta_coefs[1]))

    # likelihood:
    mu = pm.Deterministic(
        'mu',
        pm.math.dot(X, beta_coefs) + 
        pm.math.dot(Z, gamma0) +
        pm.math.dot(Z, gamma1) +
        np.log(DaysOfCatch)
    )

    # This set-up would be an over-dispersed poisson:
    # log_rate = pm.Normal('log_rate', mu=mu, sigma=alpha, shape=n_obs)
    # predy = pm.Poisson('predy', mu=pm.math.exp(log_rate), observed=y)

    # This is the neg binomial set-up:
    pm.NegativeBinomial('predy', mu=pm.math.exp(mu), alpha = alpha, observed=y)
    
    idata_pymc2nb = pm.sample(nuts={'target_accept':0.9}, draws=1000, tune=3000, cores=4, random_seed=45)

import pickle # to save PyMC object to file.
filename = 'output/idata_pymc2nb.pkl'
# save bambi object to file.
with open(filename, 'wb') as file:
    pickle.dump(idata_pymc2nb, file)


```

```{python nbPyMCOutput3}

# Load the object from the file
import pickle
filename = 'output/idata_pymc3nb.pkl'
with open(filename, 'rb') as file:
    idata_pymc3nb = pickle.load(file)

# summary stats
print(az.summary(idata_pymc3nb, var_names=["beta_coefs", "alpha", "sigma0", "sigma1"])) 

# quick diagnostics
# az.plot_trace(idata_pymc3nb, backend_kwargs={"layout": "constrained"}, legend=True, var_names=["beta_coefs", "alpha", "sigma0", "sigma1"])
# plt.show()

# az.plot_trace(idata_pymc3nb, backend_kwargs={"layout": "constrained"}, legend=True, var_names=["gamma0", "gamma1"])
# plt.show() # many are bi-model--- not a good fit!!!

# CALC TR!
print(az.summary(idata_pymc3nb, var_names=["TR"])) 

```


## GLMM with Gaussian Processes: RE for Trial, GP for Date

Now we are in the territory where we have to fit the models in a Bayesian framework. If our response variable was continuous (i.e., the regression model was based on a Normal distribution), then we could still use maximum likelihood estimation.

### Gaussian Process Bayes (brms) version

* Use numeric version of sampling date for better algorithm stability. Still, we have a warning about a divergent transition. Ideally, we would tweak the model and look at the data carefully so fix this warning, but because we are working with made up data, we will not worry abotu the warning for now.

```{r, eval =F}
summary(datR$numericDate) # check max before fitting, in case seasons don't line up
brm4nb <- brm(formula = nYSB ~ TreatmentF + offset(log(DaysOfCatch)) + 
                    (1 + TreatmentF||Location) + 
              gp(numericDate, by =Location), 
                  control = list(adapt_delta = 0.9),
              data = datR, 
              family = negbinomial,
              prior = c(set_prior("normal(0,2)", class = "b"),
                        set_prior("cauchy(0,1)", class = "shape"), # == half-cauchy
                        set_prior("normal(0,2)", class = "Intercept"),
                        set_prior("cauchy(0,1)", class = "sd"),
                        set_prior("cauchy(0,1)", class = "lscale", 
                                  coef = "gpnumericDateLocationLoc1"),
                        set_prior("cauchy(0,1)", class = "lscale", 
                                  coef = "gpnumericDateLocationLoc2"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc3"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc4"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc5"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc6"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc7"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc8"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc9"),
                        set_prior("cauchy(0,1)", class = "lscale", coef = "gpnumericDateLocationLoc10"),
                        set_prior("normal(0,2)", class = "sdgp")),
                        # set_prior("lkj(2)", class = "cor")),
              warmup = 1000, iter = 5000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(brm4nb, "../output/brm4nb.RDS")
```
```{r}
brm4nb <- readRDS("output/brm4nb.RDS")
summary(brm4nb)
# bayes_R2(brm4nb) # 0.824

# prior_summary(brm4nb)
# prior_summary(brm4nb, all = FALSE)
# 
```

```{r}
library(ggmcmc) # ggs()
# CALC TR
modtranformed <- ggs(brm4nb) # transform MCMC output into a table
beta1 <- modtranformed %>% 
  filter(Parameter == "b_TreatmentFtrtA", # our beta1
         Iteration > 500) # remove burn-in
TR = 100 * (1 - exp(beta1$value))
bayesTR_4nb <- data.frame(TR = round(median(TR), 1), 
           lowerCI = round(quantile(TR, 0.025), 1),
          upperCI = round(quantile(TR, 0.975), 1))   
knitr::kable(bayesTR_4nb)

# plot predictions:
library(tidybayes) # add_predicted_draws()
predsbrm <- fitted(brm4nb,
                   scale = "linear",
                   newdata = datR  %>% mutate(DaysOfCatch = 1))
# dim(exp(predsbrm))
# head(predsbrm)
# I used linear scale (link response) in the fitted values and then exponeniate
# because it intuitively is easier to think about moths per day (i.e., response scale),
# but I want the resolution that is had at the linear scale.
# (If I chose response scale, Id' get lots of 0's which aren't informative. 
# Give it a try!)
tmp <- cbind(datR, exp(predsbrm))
bayes_p4nb <- ggplot(tmp, 
       aes(x = DATI, y = Estimate, color = TreatmentF)) +  
  geom_line() + 
  geom_ribbon(aes(ymin = `Q2.5`, ymax = `Q97.5`, fill = TreatmentF),  # regression line and CI
                    alpha = 0.3) +
  geom_point(data = mean_cts, 
             aes(DATI, mean_mothsperday, color = TreatmentF), 
             size = 1.1) +
  facet_wrap(~Location, scales = "free_y") + 
  scale_color_manual(values = mycolors[]) +
  scale_fill_manual(values = mycolors[]) +
  labs(title = "Bayesian neg binomial GP GLMM predictions",
       subtitle = "Lines are predictions, points are the data")
print(bayes_p4nb)

```

## Summary of estimates

### Compare TR estimates

Here, the trapping reduction estimates from all of the models are compared. As expected, the median TR estimates are very close from model to model, but the uncertainty in that estimate increase (i.e., the confidence/credible intervals get wider) when we properly account for the correlations in our data.

```{r}

# Framework = Frequentist, Bayes
# Distribtuion = Poisson, NB
# Model name: GLM, GLMM (Location), GLMM (Location, nested dates), GLMM (Location, GP for correlated dates)
# Model description?
# TR
# lower CI (2.5%)
# upper CI (97.5%)
tableTR = bind_rows(
data.frame(Framework = "Frequentist", Distribution = "Poisson", Model_name = "GLM",
      round(TR_1p)),
data.frame(Framework = "Frequentist", Distribution = "Neg Binomial", Model_name = "GLM",
      round(TR_1nb)),
data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLM",
      round(bayesTR_1nb)),

data.frame(Framework = "Frequentist", Distribution = "Poisson", Model_name = "GLMM (Location)",
      round(TR_2p)),
data.frame(Framework = "Frequentist", Distribution = "Neg Binomial", Model_name = "GLMM (Location)",
      round(TR_2nb)),
data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLMM (Location)",
      round(bayesTR_2nb)),

data.frame(Framework = "Frequentist", Distribution = "Poisson", Model_name = "GLMM (Location, nested Date)",
      round(TR_3p)),
data.frame(Framework = "Frequentist", Distribution = "Neg Binomial", Model_name = "GLMM (Location, nested Date)",
      round(TR_3nb)),
data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLMM (Location, nested Date)",
      round(bayesTR_3nb)),

data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLMM (Location, Gaussian Process Date)",
      round(bayesTR_4nb))
)
rownames(tableTR) <- NULL
knitr::kable(tableTR)
```

### Compare coefficients

Because TR is a non-linear function of $\beta_1$, the differences int he model output get slightly distorted from the transformation. Thereofre, the $\beta_1$ coefficients are displayed as well:

```{r}

# Framework = Frequentist, Bayes
# Distribtuion = Poisson, NB
# Model name: GLM, GLMM (Location), GLMM (Location, nested dates), GLMM (Location, GP for correlated dates)
# Model description?
# TR
# lower CI (2.5%)
# upper CI (97.5%)
tableBeta = bind_rows(
data.frame(Framework = "Frequentist", Distribution = "Poisson", 
      Model_name = "GLM",
      Estimate = round(summary(mod1p)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod1p)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Frequentist", Distribution = "Neg Binomial",
           Model_name = "GLM",
      Estimate = round(summary(mod1nb)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod1nb)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", 
           Model_name = "GLM",
           Estimate = round(summary(brm1nb)$fixed["TreatmenttrtA", "Estimate"], 2),
      SE = round(summary(brm1nb)$fixed["TreatmenttrtA", "Est.Error"], 3)),


data.frame(Framework = "Frequentist", Distribution = "Poisson", 
           Model_name = "GLMM (Location)",
      Estimate = round(summary(mod2p)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod2p)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Frequentist", Distribution = "Neg Binomial",
           Model_name = "GLMM (Location)",
      Estimate = round(summary(mod2nb)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod2nb)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", 
           Model_name = "GLMM (Location)",
           Estimate = round(summary(brm2nb)$fixed["TreatmentFtrtA", "Estimate"], 2),
      SE = round(summary(brm2nb)$fixed["TreatmentFtrtA", "Est.Error"], 3)),


data.frame(Framework = "Frequentist", Distribution = "Poisson", 
           Model_name = "GLMM (Location, nested Date)",
      Estimate = round(summary(mod3p)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod3p)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Frequentist", Distribution = "Neg Binomial",
           Model_name = "GLMM (Location, nested Date)",
      Estimate = round(summary(mod3nb)$coefficients["TreatmentFtrt A", "Estimate"], 2),
      SE = round(summary(mod3nb)$coefficients["TreatmentFtrt A", "Std. Error"], 3)),

data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLMM (Location, nested Date)",
           Estimate = round(summary(brm3nb)$fixed["TreatmentFtrtA", "Estimate"], 2),
      SE = round(summary(brm3nb)$fixed["TreatmentFtrtA", "Est.Error"], 3)),

data.frame(Framework = "Bayesian", Distribution = "Neg Binomial", Model_name = "GLMM (Location, Gaussian Process Date)",
           Estimate = round(summary(brm4nb)$fixed["TreatmentFtrtA", "Estimate"], 2),
      SE = round(summary(brm4nb)$fixed["TreatmentFtrtA", "Est.Error"], 3))
)

rownames(tableBeta) <- NULL
knitr::kable(tableBeta)
```



### Plot comparisons

#### GLM's

In the side-by-side comparison, you can see the wider confidence intervals for the negative binomial distribution, compared to the Poisson.

```{r all_plots1}
cowplot::plot_grid(p1p_glm + theme(legend.position="none"), 
                   p1nb_glm + theme(legend.position="none"))
```

Side-by-side comparison of the frequentist vs Bayes negative binomial GLM predictions.

```{r all_plots1b}
cowplot::plot_grid(p1nb_glm + theme(legend.position="none"),
                   bayes_p1nb + theme(legend.position="none"))
```

#### GLMM's (Location RE)

Side-by-side comparison of the frequentist Poisson vs negative  binomial GLMM (Location RE) predictions.

```{r all_plots2}
cowplot::plot_grid(p2p_glmm + theme(legend.position="none"), 
                   p2nb_glmm + theme(legend.position="none"))
```

Side-by-side comparison of the frequentist vs Bayes negative binomial GLMM (Location RE) predictions.

```{r all_plots2b}
cowplot::plot_grid(p2nb_glmm + theme(legend.position="none"),
                   bayes_p2nb + theme(legend.position="none"))
```


#### GLMM's (Location RE, nested Date)

Note: These models are a great fit to the data that we have! But, they are overly confident and will not be good at predicting at new locations and new sampling dates.

Side-by-side comparison of the frequentist Poisson vs negative  binomial GLMM (Location RE) predictions.

```{r all_plots3}
cowplot::plot_grid(p3p_glmm + theme(legend.position="none"), 
                   p3nb_glmm + theme(legend.position="none"))
```

Side-by-side comparison of the frequentist vs Bayes negative binomial GLMM (Location RE) predictions.

```{r all_plots3b}
cowplot::plot_grid(p3nb_glmm + theme(legend.position="none"),
                   bayes_p3nb + theme(legend.position="none"))
```


#### GLMM's: nested Date vs Gaussian Process Date

Once we properly account for the correlations in our sampling dates, we have much more uncertainty in our overall TR predictions (see table above) , and the predicted counts are "smoothed" over time for some of the locations.

```{r all_plots4}
cowplot::plot_grid(bayes_p3nb + theme(legend.position="none"),
                   bayes_p4nb + theme(legend.position="none"))
```

