---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Nonlinear Bayesian Models in R {#NonlinearR}

(Same set up code as Chapter 3):

```{r setup, include=FALSE, message=F, warning=FALSE, results = "hide", echo=F}
knitr::opts_chunk$set(cache = T, message =F, warning = F, echo=F,
                      # dpi = 36, out.width="600px", out.height="600px")
  fig.width = 8,
  fig.height = 5,
  out.width = '100%')

# R libraries:
# install.packages("kableExtra") ## uncomment to download any required packages
library(MASS)
library(lme4)
library(magrittr)
library(tidyverse)
library(ggmap)
# usethis::edit_r_environ()
register_google(Sys.getenv("ggmapKey"))

# to run Bayesian models. Set appropriate nCores for your machine
library(brms)
nCores = 6 # nCores -2
library(ggmcmc) # ggs()
library(truncnorm)
library(rjags)
library(kableExtra)

## to use jags, you first need to download. For a Mac, go to terminal:
# brew install jags
# instructions from: https://gist.github.com/casallas/8411082

## Set colors:
# library(RColorBrewer)
# display.brewer.pal(11, "Spectral")
# brewer.pal(11, "Spectral")
mycolors = c("#9E0142", "#66C2A5")
# mycolors =  c("#9E0142", "#762A83","#9970AB",  # "#5E4FA2",
#               "#3288BD","#66C2A5", "#ABDDA4", "#D53E4F")
visualcols = c( "#D7191C", "#FDAE61", "#FFFFBF", "#A6D96A", "#1A9641")
trialcols = c("#4575B4", "#40004B","#762A83","#9970AB",  "#5AAE61",
               "#C2A5CF","#74ADD1", "#E7D4E8", "#ABD9E9",   "#00441B" )
prop_colors = c("#313695", "#ABD9E9", "#FFFFBF", "#FEE090", 
                 "#FDAE61", "#F46D43","#D73027", "#A50026")
#
```
``` {r functions}
# calculate trapping reduction for frequentist model fits.
getTR <- function (mod, coefs = -1) {
  b1 = summary(mod)$coef[coefs,1]
  seb1 = summary(mod)$coef[coefs,2]
  (TR = round(100 * (1- exp(b1)),1))
  (higher = round(100 * (1 - exp(b1 - 1.96*seb1)),1))
  (lower = round(100 * (1 - exp(b1 + 1.96*seb1)),1)) 
  data.frame(#Treatment = names(TR),
             TR = as.numeric(TR), 
             lowerCI = as.numeric(lower),
             upperCI = as.numeric(higher))   
}

## NA = 0 in sum
`%+%` <- function(x, y)  mapply(sum, x, y, MoreArgs = list(na.rm = TRUE))
#

## prediction plot for frequentist models:
predplot <- function(preddata, plottitle = "") {
  p = ggplot(preddata,
       aes(DATI, preds, color = TreatmentF)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowerCI, ymax = upperCI, fill = TreatmentF),
              alpha = 0.3) + 
  geom_point(data = mean_cts,
             aes(DATI, mean_mothsperday, color = TreatmentF), size = 1.1) +
  facet_wrap(~Location, scales = "free_y") + 
  scale_color_manual(values = mycolors[]) +
  scale_fill_manual(values = mycolors[]) +
  labs(title = plottitle, 
       subtitle = "Lines are predictions, points are the data")

  return(p)
}

brmspredplot = function(preddata, plottitle = "") {
  p <- ggplot(preddata, 
       aes(x = DATI, y = Estimate, color = TreatmentF)) +  
  geom_line() + 
  geom_ribbon(aes(ymin = `Q2.5`, ymax = `Q97.5`, 
                  fill = TreatmentF),  # regression line and CI
                    alpha = 0.3) +
  geom_point(data = mean_cts, 
             aes(DATI, mean_mothsperday, color = TreatmentF), 
             size = 1.1) +
  facet_wrap(~Location, scales = "free_y") + 
  scale_color_manual(values = mycolors[]) +
  scale_fill_manual(values = mycolors[]) +
  labs(title = plottitle,
       subtitle = "Lines are predictions, points are the data")

  return(p)
}

```
```{r readdatR}

datRinit = read.csv("data/moths.csv")
# drop first column, is just unnecessary indexing:
datRinit %<>% dplyr::select(-X)
# (there is a select function in MASS package that sometimes overrides dplyr)
```
```{r data_manip, results = "hide"}
# check data values: 
str(datRinit)
summary(datRinit)
lapply(datRinit, function(x) {
  if (is.character(x))
    unique(x)
})
# convert dates to dates in case further manipulation on them is needed:
datR <- datRinit %>%
  mutate(TransplantDate = ymd(TransplantDate),
         DispInstallDate = ymd(DispInstallDate),
         TrapInstallDate = ymd(TrapInstallDate),
         SamplingDate = ymd(SamplingDate)
         )  %>%
  mutate(mothsperday = nYSB / DaysOfCatch, # standardize the counts
         TreatmentF = factor(Treatment), # if we need to fix levels in non-alpha order
         LocationF = as.factor(Location),
         # DATscaled = as.numeric(scale(DAT)),
         SamplingDateC = as.character(SamplingDate),
         numericDate = as.numeric(SamplingDate),
         numericDate = numericDate - min(numericDate) + 1)

datR %<>%
  mutate(numTrt = ifelse(Treatment == "CGP", 0, 1))

## average counts for each location, date
mean_cts <- datR %>%
  group_by(Location, Treatment, TreatmentF,
           AssessmentNumber,
           SamplingDate, DATI) %>%
  summarize(mean_cts = mean(nYSB, na.rm = T),
            mean_mothsperday = mean(mothsperday)) %>%
  ungroup()


```

Asymptotic sigmoidal curve expected for product performance over time. Eventually, given enough time, the product has to stop working, and $\beta_1$ goes to 0 as t goes to 0.

## Generalized Logistic Curve

Adapted from: https://en.wikipedia.org/wiki/Generalised_logistic_function

$$
\beta(t) = A + \frac{(K - A)}{(1 + e^{-B(t-M)})^{1/v}}
$$
where,

$A$ is the left horizontal asymptote,

$K$ is the right horizontal asymptote ($K = 0$ in our models),

$B$ is the growth rate ($B>0$; $B<0$ is a decay rate),

$v$ affects where the inflection point occurs, and

$M$ relates to the starting time of when the curve begins.

```{r SimData}

A = -5 # left asymptote
K = 0 # right asymptote
B = 0.2 # (decay rate if b < 0; growth if b > 0)
v = 1 # affects where inflection point is
# Q = 9 # related to y at time 0 (how long until y starts to go to 0)
m = 25 # expected DAI when TR failure is midway
Q = exp(B*m)
Q
C = 1

t = 0:150
beta = A + ( (K - A)  / ( (C + Q * exp(-1 * t * B)) ^ (1 / v) ) )
TR = 100 * (1 - exp(beta))
plot(t, beta, type = "l")
# plot(t, TR, type = "l")

```

See the shiny app to play around with how the parameter values affect the curve. This will be helpful in understanding the model and setting priors for the model.

https://bromsk.shinyapps.io/GeneralizedLogisticCurves/


## Bayesian model

$$
y_{ijk} \sim NegBinom(\lambda_{ijk}, r) \\
log(\lambda_{ikt} ) = \beta_0 + \beta_{1,t} x_{A,i} + \gamma_{k} x_{ik}  + GP_{ijk}  + \mbox{ln} \left(DaysOfCatch_i\right) \\
\beta_{1,t} = A_k \left(1 - \frac{1}{(1 + e^{-B(t-M)})^{1/v}}\right) \\
\gamma_k \sim Normal(0, \sigma_{\gamma}) \\ 
\mathbf{GP_k} \sim MVN(0, \boldsymbol\Sigma_k) \\
\beta_0 \sim Normal(0, 2) \\
A_k \sim HalfNormal(0, 2) \\
\sigma_{\gamma} \sim HalfCauchy(0, 1) \\
r \sim HalfCauchy(0,1) \\
B \sim HalfNormal(0.5) \\
M \sim Uniform(0, 150) \\
\sigma_{k, lm} = \tau_k \cdot e^{-\frac{|d_l - d_m|^2}{l_k^2}} \\
\tau_k \sim HalfCauchy(0, 1) \\
l_k \sim HalfCauchy(0, 1) \\
$$

where, in addition to the variables and parameters defined for the GLM, we have

$k = 1,..., K=10$ represents the locations, 

$j = 1, ..., J = 117$ represents the 117 unique date-location combinations,

$y_{ikj}$ is moth count $i$ from location $k$ on date $j$,

$\lambda_{ikj}$ is the expected moth count for sample $i$  from location $k$ on date $j$,

$\beta_0$ is the expected moth count for our control treatment for a new location,

$\beta_{1,t}$ is the expected difference in moth counts between the control group and the treatment A group for a new location at $t$ days from the installation, 

$A_k$ is the maximum trapping reduction (represented by the max coefficient value) associated with locaiton $k$,

$\gamma_{0k}$ is the random intercept associated with location $k$, which leads to different background moth pressures at each location. All $\gamma_{0k}$ come from an iid Normal distribution,

$\gamma_{1k}$ is the random slope associated with location $k$, which leads to different treatment effects at each location. All $\gamma_{1k}$ come from an iid Normal distribution, 

$x_{ik}$ is an indicator variable that equals 1 if moth count $i$ is associated with location $k$ and 0 otherwise,

$\gamma_{00j}$ is the random intercept associated with date-location $j$, which leads to different background moth pressures for each date, and 

$x_{ij}$ is an indicator variable that equals 1 if moth count $i$ is associated with date-location $j$ and 0 otherwise.

$l$ = lscale (differnt for each location)

$\phi$ = sdgp

For now, we assume that each location shares all of the same non-linear parameters except for $A$, the maximum trapping reduction level. For now, we also set $v=1$.

## Fitting models with the brms package

Setting priors in a nonlinear model is non-trivial!!

### no RE, no GP

```{r simdata1}
DAT = 11:114
DATsampled = seq(11, 114, by = 7)
Treatment = c("CGP", "trt A")
Location = c("A", "B", "C", "D", "E")
Trap = 1:4
# simdata1 = expand_grid(Location, Treatment, Trap, DAT= DATsampled)

b0 = 2
A = -3
B = 0.1
M = 50
b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))

simdata1 <- expand_grid(Location, Treatment, Trap, DAT= DATsampled) %>%
  mutate(DaysOfCatch = 1,
         numTrt = ifelse(Treatment == "CGP", 0, 1))
simdata1 %<>%
  mutate(b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) ))),
         TR = 100 * (1 - exp(b1)),
         muTmp = exp(b0 + b1*numTrt),
         nYSB = rnbinom(n=nrow(simdata1), mu = muTmp, size = 5))
# head(simdata)
# nYSB = rnbinom(n=length(DAT), mu = exp(b0 + b1), size = 5)
# var = mu + mu^2 /size
# nYSB = rpois(n=length(DAT), lambda = exp(b0 + b1))
# plot(DAT, nYSB, type = "l")

ggplot(simdata1, aes(DAT, nYSB, color = Treatment)) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~Location)

```
```{r fit1, eval =F}
prior1 <- prior(normal(0,2), nlpar = "b0") +
  prior(normal(0, 2), lb = 0, nlpar = "A") + 
  prior(normal(0, 0.5), lb = 0, nlpar = "B") +
  prior(uniform(0, 120), lb = 0, ub = 120, nlpar = "M") +
  prior(cauchy(0,1), lb = 0, class = "shape")

fit1 <- brm(bf(nYSB ~ b0 - (A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))) * numTrt + 
                 #(1|Location) + 
              # gp(DAT, by =Location)  + 
                log(DaysOfCatch), # offset() is implicit in nonlinear formula
              b0 + A + B + M ~ 1,
              nl = TRUE), 
              control = list(adapt_delta = 0.9),
              data = simdata1, 
              family = negbinomial,
              prior = prior1,
                        # set_prior("cauchy(0,1)", class = "shape"), 
              warmup = 1000, iter = 2000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(fit1, "output/fit1.RDS")
```

```{r summary1}
fit1 <- readRDS("output/fit1.RDS")
summary(fit1)
# plot(fit1)
prior_summary(fit1)
# 
```

```{r output1}
library(ggmcmc) # ggs()
# CALC TR
modtranformed <- ggs(fit1) # transform MCMC output into a table
# str(modtranformed)
A1 <- modtranformed %>% 
  filter(Parameter == c("b_A_Intercept"), 
         Iteration > 500) # remove burn-in
B1 <- modtranformed %>% 
  filter(Parameter == c("b_B_Intercept"),
         Iteration > 500) # remove burn-in
M1 <- modtranformed %>% 
  filter(Parameter == c("b_M_Intercept"),
         Iteration > 500) # remove burn-in
DATmatrix = matrix(DAT, nrow = nrow(A1), ncol = length(DAT), byrow = T)
tmp = apply(DATmatrix, 2, function(x) x - M1$value) # nIter x nDAT
beta1 = apply(tmp, 2, function(x) -1 * A1$value * (1 - 1 / (1 + exp(-1 * B1$value * x ))))
TR = 100 * (1 - exp(beta1))

TR1 <- data.frame(TR = round(apply(TR, 2, median), 1), 
           lowerCI = round(apply(TR, 2, quantile, 0.025), 1), 
          upperCI = round(apply(TR, 2, quantile, 0.975), 1),
          DAT = DAT)
ggplot(TR1) +
  geom_line(aes(DAT, TR), size = 1.2) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI), alpha = 0.3) +
  geom_line(data = simdata1, aes(DAT, TR), color = "purple", size = 1.2) +
  labs(title = "Predicted and true TR curve")

```


### intercept RE, no GP

```{r simdata2}
DAT = 11:114
DATsampled = seq(11, 114, by = 7)
Treatment = c("CGP", "trt A")
Location = c("A", "B", "C", "D", "E")
Trap = 1:4

set.seed(0528)
b0 = rnorm(n = length(Location), 2, sd = 0.8) # now have 5 intercepts for 5 loc's
A = -3
B = 0.1
M = 50
b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))

simdata2 = expand_grid(Location, Treatment, Trap, DAT= DATsampled) %>%
  mutate(DaysOfCatch = 1,
         numTrt = ifelse(Treatment == "CGP", 0, 1)) %>%
  left_join(data.frame(b0, Location))
simdata2 %<>%
  mutate(b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) ))),
         muTmp = exp(b0 + b1*numTrt),
         nYSB = rnbinom(n=nrow(simdata2), mu = muTmp, size = 5))
ggplot(simdata2, aes(DAT, nYSB, color = Treatment)) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~Location)

```
```{r, eval =F}
prior2 <- prior(normal(0,2), nlpar = "b0") +
  prior(normal(0, 2), lb = 0, nlpar = "A") + 
  prior(normal(0, 0.5), lb = 0, nlpar = "B") +
  prior(uniform(0, 120), lb = 0, ub = 120, nlpar = "M")

fit2 <- brm(bf(nYSB ~ b0 - (A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))) * numTrt + 
              # gp(DAT, by =Location)  + 
                log(DaysOfCatch), # offset() is implicit in nonlinear formula
              b0 ~ 1 + (1|Location),
              A + B + M ~ 1,
              nl = TRUE), 
              control = list(adapt_delta = 0.9),
              data = simdata2, 
              family = poisson,
              prior = prior2,
                        # set_prior("cauchy(0,1)", class = "shape"), 
              warmup = 1000, iter = 2000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(fit2, "output/fit2.RDS")
```

```{r}
fit2 <- readRDS("output/fit2.RDS")
summary(fit2)
# plot(fit2)
# conditional_effects(fit2, "DAT:numTrt")
# # 
# conditions <- data.frame(Location = unique(simdata$Location))
# rownames(conditions) <- unique(simdata$Location)
# me_loss <- conditional_effects(
#   fit2, conditions = conditions,
#   re_formula = NULL, method = "predict"
# )
# plot(me_loss, ncol = 5, points = TRUE)

```

```{r output2}
library(ggmcmc) # ggs()
# CALC TR
modtranformed <- ggs(fit2) # transform MCMC output into a table
# get b0 estimates for each location:
meanb0est <- modtranformed %>%
  filter(Parameter == "b_b0_Intercept") %>%
  summarize(med = median(value))
modtranformed %>%
  filter(grepl("r_", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(med = median(value) + meanb0est$med)
round(b0, 2)
# str(modtranformed)
estA <- modtranformed %>% 
  filter(Parameter == c("b_A_Intercept"), 
         Iteration > 500) # remove burn-in
estB <- modtranformed %>% 
  filter(Parameter == c("b_B_Intercept"),
         Iteration > 500) # remove burn-in
estM <- modtranformed %>% 
  filter(Parameter == c("b_M_Intercept"),
         Iteration > 500) # remove burn-in
DATmatrix = matrix(DAT, nrow = nrow(estA), ncol = length(DAT), byrow = T)
tmp = apply(DATmatrix, 2, function(x) x - estM$value) # nIter x nDAT
beta1 = apply(tmp, 2, function(x) -1 * estA$value * (1 - 1 / (1 + exp(-1 * estB$value * x ))))
TR = 100 * (1 - exp(beta1))

TRdf <- data.frame(TR = round(apply(TR, 2, median), 1), 
           lowerCI = round(apply(TR, 2, quantile, 0.025), 1), 
          upperCI = round(apply(TR, 2, quantile, 0.975), 1),
          DAT = DAT)
ggplot(TRdf) +
  geom_line(aes(DAT, TR), size = 1.2) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI), alpha = 0.3) +
  geom_line(data = simdata1, aes(DAT, TR), color = "purple", size = 1.2) +
  labs(title = "Predicted and true TR curve")

```

### intercept RE, A RE, no GP

```{r simdata3}
DAT = 11:114
DATsampled = seq(11, 114, by = 7)
Treatment = c("CGP", "trt A")
Location = c("A", "B", "C", "D", "E")
Trap = 1:4
# simdata = expand_grid(Location, Treatment, Trap, DAT= DATsampled)

set.seed(0528)
b0 = rnorm(n = length(Location), 2, sd = 0.8) # now have 5 intercepts for 5 loc's
A = rnorm(n = length(Location), -3, sd = 0.5)
B = 0.1
M = 50
b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))

simdata3 <- expand_grid(Location, Treatment, Trap, DAT= DATsampled) %>%
  mutate(DaysOfCatch = 1,
         numTrt = ifelse(Treatment == "CGP", 0, 1)) %>%
  left_join(data.frame(b0, Location))  %>%
  left_join(data.frame(A, Location))

simdata3 %<>%
  mutate(b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) ))),
         muTmp = exp(b0 + b1*numTrt),
         nYSB = rnbinom(n=nrow(simdata3), mu = muTmp, size = 5))

ggplot(simdata3, aes(DAT, nYSB, color = Treatment)) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~Location)

```
```{r fit3, eval =F}
prior3 <- prior(normal(0,2), nlpar = "b0") +
  prior(normal(0, 2), lb = 0, nlpar = "A") + 
  prior(normal(0, 0.5), lb = 0, nlpar = "B") +
  prior(uniform(0, 120), lb = 0, ub = 120, nlpar = "M") +
  prior(cauchy(0,1), lb = 0, class = "shape")

fit3 <- brm(bf(nYSB ~ b0 - (A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))) * numTrt + 
              # gp(DAT, by =Location)  + 
                log(DaysOfCatch), # offset() is implicit in nonlinear formula
              b0 ~ 1 + (1|Location),
              A  ~ 1 + (1|Location),
              B + M ~ 1,
              nl = TRUE), 
              control = list(adapt_delta = 0.9),
              data = simdata3, 
              family = negbinomial,
              prior = prior3,
              warmup = 1000, iter = 2000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(fit3, "output/fit3.RDS")
```
```{r summary3}
fit3 <- readRDS("output/fit3.RDS")
summary(fit3)
# plot(fit3)
# conditional_effects(fit3, "DAT:numTrt")
# # 

```

```{r output3}
library(ggmcmc) # ggs()
# CALC TR
modtranformed <- ggs(fit3) # transform MCMC output into a table
# get b0 estimates for each location:
meanb0est <- modtranformed %>%
  filter(Parameter == "b_b0_Intercept") %>%
  summarize(med = median(value))
modtranformed %>%
  filter(grepl("r_Location__b0", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(med = median(value) + meanb0est$med)
round(b0, 2)
# str(modtranformed)
# look at location-specific A estimates:
estA <- modtranformed %>% 
  filter(Parameter == c("b_A_Intercept"), 
         Iteration > 500) 
modtranformed %>%
  filter(grepl("r_Location__A", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(meddiff = median(value)) %>%
  mutate(med = meddiff + median(estA$value))
round(-1*A, 2)

# calculate overall avg TR for new location:
estB <- modtranformed %>% 
  filter(Parameter == c("b_B_Intercept"),
         Iteration > 500) # remove burn-in
estM <- modtranformed %>% 
  filter(Parameter == c("b_M_Intercept"),
         Iteration > 500) # remove burn-in
DATmatrix = matrix(DAT, nrow = nrow(estA), ncol = length(DAT), byrow = T)
tmp = apply(DATmatrix, 2, function(x) x - estM$value) # nIter x nDAT
beta1 = apply(tmp, 2, function(x) -1 * estA$value * (1 - 1 / (1 + exp(-1 * estB$value * x ))))
TR = 100 * (1 - exp(beta1))

TRdf <- data.frame(TR = round(apply(TR, 2, median), 1), 
           lowerCI = round(apply(TR, 2, quantile, 0.025), 1), 
          upperCI = round(apply(TR, 2, quantile, 0.975), 1),
          DAT = DAT)
ggplot(TRdf) +
  geom_line(aes(DAT, TR), size = 1.2) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI), alpha = 0.3) +
  geom_line(data = simdata1, aes(DAT, TR), color = "purple", size = 1.2) +
  labs(title = "Predicted and true TR curve")

```
```{r preds3}
## predict moth counts for each location.
preds = posterior_predict(fit3)
str(preds) # nIter x nrow(simdata)
predsDF <- bind_cols(simdata3,
                     preds = apply(preds, 2, median),
                     lowerCI = apply(preds, 2, quantile, 0.025),
                     upperCI = apply(preds, 2, quantile, 0.975))
ggplot(predsDF) +
  geom_line(aes(DAT, preds, color = Treatment)) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI, fill = Treatment), alpha = 0.3) +
  geom_point(data = simdata3, aes(DAT, nYSB, color = Treatment)) +
  facet_wrap(~Location)
```

### intercept RE, A RE, WITH GP (same for all loc)

```{r simdata4}
# simulate the GP:
x = DATsampled # runif(100)
d = abs(outer(x, x, "-")) # compute distance matrix, d_{ij} = |x_i - x_j|
l = 30 # length scale- smaller l is more wiggle
tau = 3
Sigma_SE = tau * exp(-d^2/(2*l^2)) # squared exponential kernel
set.seed(100)
for (i in 1:length(Location)) {
  GP = as.vector(mvtnorm::rmvnorm(1,sigma=Sigma_SE))
  plot(DATsampled, GP, type = "l")
  # if (all(GP<0)) GP = abs(GP)
  # GP = abs(GP)
  if (i == 1) {
   out = data.frame(DAT = DATsampled,
             GP = as.vector(GP),
             Location = Location[i])
  } else {
    out %<>%
      bind_rows(data.frame(DAT = DATsampled,
             GP = as.vector(GP),
             Location = Location[i]))
  }
}
#

DAT = 11:114
DATsampled = seq(11, 114, by = 7)
Treatment = c("CGP", "trt A")
Location = c("A", "B", "C", "D", "E")
Trap = 1:4

# str(dat)
set.seed(0530)
b0 = rnorm(n = length(Location), 1, sd = 1.5) # now have 5 intercepts for 5 loc's
A = rnorm(n = length(Location), -3, sd = 0.5)
B = 0.1
M = 50
# b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))

simdata4 <- expand_grid(Location, Treatment, Trap, DAT= DATsampled) %>%
  mutate(DaysOfCatch = 1,
         numTrt = ifelse(Treatment == "CGP", 0, 1)) %>%
  left_join(data.frame(b0, Location))  %>%
  left_join(data.frame(A, Location)) %>%
  left_join(out)
# head(simdata4)
simdata4 %<>%
  mutate(b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) ))),
         muTmp = exp(b0 + b1*numTrt + GP),
         nYSB = rnbinom(n=nrow(simdata4), mu = muTmp, size = 30))
print(sum(is.na(simdata4$nYSB)))
simdata4 %<>%
  mutate(nYSB = ifelse(is.na(nYSB), 0, nYSB))
# hist(simdata4$nYSB)
ggplot(simdata4, aes(DAT, nYSB, color = Treatment)) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~Location, scales = "free")

```
```{r fit4, eval =F}
# https://bookdown.org/content/4857/adventures-in-covariance.html#continuous-categories-and-the-gaussian-process
prior4 <- prior(normal(0,2), nlpar = "b0") +
  prior(normal(0, 2), lb = 0, nlpar = "A") + 
  prior(normal(0, 0.5), lb = 0, nlpar = "B") +
  prior(uniform(0, 120), lb = 0, ub = 120, nlpar = "M") +
  prior(cauchy(0,1), lb = 0, class = "shape") +
  prior(cauchy(0,50), lb = 0, class = "lscale", nlpar = "gpP") +
  prior(cauchy(0,1), lb = 0, class = "sdgp", nlpar = "gpP")

fit4 <- brm(bf(nYSB ~  b0 - (A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))) * numTrt + 
              gpP + 
              log(DaysOfCatch), 
              b0 ~ (1|Location),
              gpP ~ 1 + gp(DAT), # GP absorbs the intercept
              A  ~ 1 + (1|Location),
              B + M ~ 1,
              nl = TRUE), 
              control = list(adapt_delta = 0.9),
              data = simdata4, 
              family = negbinomial,
              prior = prior4,
              warmup = 1000, iter = 2000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(fit4, "output/fit4.RDS")
```
```{r summary4}
fit4 <- readRDS("output/fit4.RDS")
summary(fit4)
# plot(fit4)
# conditional_effects(fit4, "DAT:numTrt")
plot(conditional_effects(fit4, "DAT:numTrt"), points = T)
# # 

```

```{r output4}
library(ggmcmc) # ggs()
# CALC TR
modtranformed <- ggs(fit4) # transform MCMC output into a table
unique(modtranformed$Parameter)
# get b0 estimates for each location:
meanb0est <- modtranformed %>%
  filter(Parameter == "b_b0_Intercept") %>%
  summarize(med = median(value))
meanGP0est <- modtranformed %>%
  filter(Parameter == "b_gpP_Intercept") %>%
  summarize(med = median(value))
modtranformed %>%
  filter(grepl("r_Location__b0", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(med = median(value) + meanb0est$med + meanGP0est$med)
round(b0, 2)
# str(modtranformed)
# look at location-specific A estimates:
estA <- modtranformed %>% 
  filter(Parameter == c("b_A_Intercept"), 
         Iteration > 500) 
modtranformed %>%
  filter(grepl("r_Location__A", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(meddiff = median(value)) %>%
  mutate(med = meddiff + median(estA$value))
round(-1*A, 2)

# calculate overall avg TR for new location:
estB <- modtranformed %>% 
  filter(Parameter == c("b_B_Intercept"),
         Iteration > 500) # remove burn-in
estM <- modtranformed %>% 
  filter(Parameter == c("b_M_Intercept"),
         Iteration > 500) # remove burn-in
DATmatrix = matrix(DAT, nrow = nrow(estA), ncol = length(DAT), byrow = T)
tmp = apply(DATmatrix, 2, function(x) x - estM$value) # nIter x nDAT
beta1 = apply(tmp, 2, function(x) -1 * estA$value * (1 - 1 / (1 + exp(-1 * estB$value * x ))))
TR = 100 * (1 - exp(beta1))

TRdf <- data.frame(TR = round(apply(TR, 2, median), 1), 
           lowerCI = round(apply(TR, 2, quantile, 0.025), 1), 
          upperCI = round(apply(TR, 2, quantile, 0.975), 1),
          DAT = DAT)
ggplot(TRdf) +
  geom_line(aes(DAT, TR), size = 1.2) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI), alpha = 0.3) +
  geom_line(data = simdata1, aes(DAT, TR), color = "purple", size = 1.2) +
  labs(title = "Predicted and true TR curve")

```
```{r preds4}

## predict moth counts for each location.
preds = posterior_predict(fit4)
str(preds) # nIter x nrow(simdata)
predsDF <- bind_cols(simdata4,
                     preds = apply(preds, 2, median),
                     lowerCI = apply(preds, 2, quantile, 0.025),
                     upperCI = apply(preds, 2, quantile, 0.975))
ggplot(predsDF) +
  geom_line(aes(DAT, preds, color = Treatment)) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI, fill = Treatment), alpha = 0.3) +
  geom_point(data = simdata4, aes(DAT, nYSB, color = Treatment)) +
  facet_wrap(~Location)
#
```


### intercept, A RE, with GP (DIFF for each loc)

```{r simdata5}
DAT = 11:114
DATsampled = seq(11, 114, by = 7)
Treatment = c("CGP", "trt A")
Location = c("A", "B", "C", "D", "E", "F", "G", "H", "I")
Trap = 1:4
nLocs = length(Location)

# simulate the GP:
x = DATsampled # runif(100)
d = abs(outer(x, x, "-")) # compute distance matrix, d_{ij} = |x_i - x_j|
l = rnorm(n = nLocs, mean = 30, sd = 10) # smaller l is more wiggle
tau = rnorm(n = nLocs, mean = 3, sd = 1)
set.seed(100)
for (i in 1:nLocs) {
  Sigma_SE = tau[i] * exp(-d^2/(2*l[i]^2)) # squared exponential kernel
  GP = as.vector(mvtnorm::rmvnorm(1,sigma=Sigma_SE))
  if (all(GP<0)) GP = abs(GP)
  # GP = abs(GP)
  plot(DATsampled, GP, type = "l")
  if (i == 1) {
   out = data.frame(DAT = DATsampled,
             GP = as.vector(GP),
             Location = Location[i])
  } else {
    out %<>%
      bind_rows(data.frame(DAT = DATsampled,
             GP = as.vector(GP),
             Location = Location[i]))
  }
}
#

set.seed(0530)
b0 = rnorm(n = nLocs, 1, sd = 1.5) # now have 5 intercepts for 5 loc's
A = rnorm(n = nLocs, -3, sd = 0.5)
B = 0.1
M = 50
# b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))

simdata5 <- expand_grid(Location, Treatment, Trap, DAT= DATsampled) %>%
  mutate(DaysOfCatch = 1,
         numTrt = ifelse(Treatment == "CGP", 0, 1)) %>%
  left_join(data.frame(b0, Location))  %>%
  left_join(data.frame(A, Location)) %>%
  left_join(out)
# head(simdata5)
simdata5 %<>%
  mutate(b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) ))),
         muTmp = exp(b0 + b1*numTrt + GP),
         nYSB = rnbinom(n=nrow(simdata5), mu = muTmp, size = 30))
print(sum(is.na(simdata5$nYSB)))
simdata5 %<>%
  mutate(nYSB = ifelse(is.na(nYSB), 0, nYSB))
# hist(simdata4$nYSB)
ggplot(simdata5, aes(DAT, nYSB, color = Treatment)) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~Location, scales = "free")

```
```{r fit5, eval =F}
# https://bookdown.org/content/4857/adventures-in-covariance.html#continuous-categories-and-the-gaussian-process
prior5 <- prior(normal(0,2), nlpar = "b0") +
  prior(normal(0, 2), lb = 0, nlpar = "A") + 
  prior(normal(0, 0.5), lb = 0, nlpar = "B") +
  prior(uniform(0, 120), lb = 0, ub = 120, nlpar = "M") +
  prior(cauchy(0,1), lb = 0, class = "shape") +
  prior(cauchy(0,50), lb = 0, class = "lscale", nlpar = "gpP") +
  prior(cauchy(0,10), lb = 0, class = "sdgp", nlpar = "gpP")

fit5 <- brm(bf(nYSB ~  b0 - (A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))) * numTrt +
              gpP + 
              log(DaysOfCatch), 
              b0 ~ (1|Location),
              gpP ~ 1 + gp(DAT, by = Location), # GP absorbs the intercept
              A  ~ 1 + (1|Location),
              B + M ~ 1,
              nl = TRUE), 
              control = list(adapt_delta = 0.9),
              data = simdata5, 
              prior = prior5,
              family = negbinomial,
              warmup = 1000, iter = 2000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(fit5, "output/fit5.RDS")
```
```{r summary5}
fit5 <- readRDS("output/fit5.RDS")
summary(fit5)
# plot(fit5)
# conditional_effects(fit4, "DAT:numTrt")
plot(conditional_effects(fit5, "DAT:numTrt"), points = T)
# # 

```

```{r output5}
library(ggmcmc) # ggs()
# CALC TR
modtranformed <- ggs(fit5) # transform MCMC output into a table
unique(modtranformed$Parameter)
# get b0 estimates for each location:
meanb0est <- modtranformed %>%
  filter(Parameter == "b_b0_Intercept") %>%
  summarize(med = median(value))
meanGP0est <- modtranformed %>%
  filter(Parameter == "b_gpP_Intercept") %>%
  summarize(med = median(value))
modtranformed %>%
  filter(grepl("r_Location__b0", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(med = median(value) + meanb0est$med + meanGP0est$med)
round(b0, 2)
# str(modtranformed)
# look at location-specific A estimates:
estA <- modtranformed %>% 
  filter(Parameter == c("b_A_Intercept"), 
         Iteration > 500) 
modtranformed %>%
  filter(grepl("r_Location__A", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(meddiff = median(value)) %>%
  mutate(med = meddiff + median(estA$value))
round(-1*A, 2)

# calculate overall avg TR for new location:
estB <- modtranformed %>% 
  filter(Parameter == c("b_B_Intercept"),
         Iteration > 500) # remove burn-in
estM <- modtranformed %>% 
  filter(Parameter == c("b_M_Intercept"),
         Iteration > 500) # remove burn-in
DATmatrix = matrix(DAT, nrow = nrow(estA), ncol = length(DAT), byrow = T)
tmp = apply(DATmatrix, 2, function(x) x - estM$value) # nIter x nDAT
beta1 = apply(tmp, 2, function(x) -1 * estA$value * (1 - 1 / (1 + exp(-1 * estB$value * x ))))
TR = 100 * (1 - exp(beta1))

TRdf <- data.frame(TR = round(apply(TR, 2, median), 1), 
           lowerCI = round(apply(TR, 2, quantile, 0.025), 1), 
          upperCI = round(apply(TR, 2, quantile, 0.975), 1),
          DAT = DAT)
ggplot(TRdf) +
  geom_line(aes(DAT, TR), size = 1.2) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI), alpha = 0.3) +
  geom_line(data = simdata1, aes(DAT, TR), color = "purple", size = 1.2) +
  labs(title = "Predicted and true TR curve")

```
```{r preds5}

## predict moth counts for each location.
preds = posterior_predict(fit5)
str(preds) # nIter x nrow(simdata)
predsDF <- bind_cols(simdata4,
                     preds = apply(preds, 2, median),
                     lowerCI = apply(preds, 2, quantile, 0.025),
                     upperCI = apply(preds, 2, quantile, 0.975))
ggplot(predsDF) +
  geom_line(aes(DAT, preds, color = Treatment)) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI, fill = Treatment), alpha = 0.3) +
  geom_point(data = simdata4, aes(DAT, nYSB, color = Treatment)) +
  facet_wrap(~Location)
#
```


### intercept, A, B, M RE, with GP (DIFF for each loc)

```{r simdata6}
set.seed(600)
DAT = 11:114
DATsampled = seq(11, 114, by = 7)
Treatment = c("CGP", "trt A")
Location = c("A", "B", "C", "D", "E", "F", "G", "H", "I")
Trap = 1:4
nLocs = length(Location)

# simulate the GP:
x = DATsampled # runif(100)
d = abs(outer(x, x, "-")) # compute distance matrix, d_{ij} = |x_i - x_j|
l = rnorm(n = nLocs, mean = 30, sd = 10) # smaller l is more wiggle
tau = rnorm(n = nLocs, mean = 1, sd = 0.5)
# set.seed(100)
for (i in 1:nLocs) {
  Sigma_SE = tau[i]^2 * exp(-d^2/(2*l[i]^2)) # squared exponential kernel
  GP = as.vector(mvtnorm::rmvnorm(1,sigma=Sigma_SE))
  if (all(GP<0)) GP = abs(GP)
  # GP = abs(GP)
  # plot(DATsampled, GP, type = "l")
  if (i == 1) {
   out = data.frame(DAT = DATsampled,
             GP = as.vector(GP),
             Location = Location[i])
  } else {
    out %<>%
      bind_rows(data.frame(DAT = DATsampled,
             GP = as.vector(GP),
             Location = Location[i]))
  }
}
#

# set.seed(0530)
b0 = rnorm(n = nLocs, 1, sd = 0.5) # now have 5 intercepts for 5 loc's
A = rnorm(n = nLocs, -3, sd = 0.5)
library(truncnorm)
B = rtruncnorm(n = nLocs,a = 0, mean = 0.1, sd = 0.02)
M = rtruncnorm(n = nLocs, a = 0, mean = 50, sd = 10)
# b1 = A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))

simdata6 <- expand_grid(Location, Treatment, Trap, DAT = DATsampled) %>%
  mutate(DaysOfCatch = 1,
         numTrt = ifelse(Treatment == "CGP", 0, 1)) %>%
  left_join(data.frame(b0, Location))  %>%
  left_join(data.frame(A, Location)) %>%
  left_join(data.frame(B, Location)) %>%
  left_join(data.frame(M, Location)) %>%
  left_join(out)
# head(simdata4)
simdata6 %<>%
  mutate(b1 = A * (1 - (1 / (1 + exp(-1 * B * (DAT - M) )))),
         # b1 = A + ((0 - A) / ( (1 + exp(-1 * B * (DAT - M))) ) ),
         muTmp = exp(b0 + b1*numTrt + GP),
         nYSB = rnbinom(n=nrow(simdata6), mu = muTmp, size = 30))
# print(sum(is.na(simdata6$nYSB)))
simdata6 %<>%
  mutate(nYSB = ifelse(is.na(nYSB), 0, nYSB))
# hist(simdata4$nYSB)
ggplot(simdata6, aes(DAT, nYSB, color = Treatment)) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~Location, scales = "free")

```
```{r fit6, eval =F}
# https://bookdown.org/content/4857/adventures-in-covariance.html#continuous-categories-and-the-gaussian-process
prior6 <- prior(normal(0,2), nlpar = "b0") +
  prior(normal(0, 2), lb = 0, nlpar = "A") + 
  prior(normal(0, 0.5), lb = 0, nlpar = "B") +
  prior(uniform(0, 120), lb = 0, ub = 120, nlpar = "M") +
  prior(cauchy(0,1), lb = 0, class = "shape") +
  prior(cauchy(0,50), lb = 0, class = "lscale", nlpar = "gpP") +
  prior(cauchy(0,10), lb = 0, class = "sdgp", nlpar = "gpP")

fit6 <- brm(bf(nYSB ~  b0 - (A * (1 - (1 / (1 + exp(-1 * B * (DAT - M) ))))) * numTrt +
              gpP + 
              log(DaysOfCatch), 
              b0 ~ 1 + (1|Location),
              gpP ~ 1 + gp(DAT, by = Location), # GP absorbs the intercept
              A  ~ 1 + (1|Location),
              B  ~ 1 + (1|Location),
              M  ~ 1 + (1|Location),
              nl = TRUE), 
              control = list(adapt_delta = 0.9),
              data = simdata6, 
              family = negbinomial,
              prior = prior6,
              warmup = 1000, iter = 2000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(fit6, "output/fit6.RDS")
```
```{r summary6}
fit6 <- readRDS("output/fit6.RDS")
summary(fit6)
# plot(fit6)
# conditional_effects(fit4, "DAT:numTrt")
plot(conditional_effects(fit6, "DAT:numTrt"), points = T)
# # 

```

```{r output6}
library(ggmcmc) # ggs()
# CALC TR
modtranformed <- ggs(fit6) # transform MCMC output into a table
unique(modtranformed$Parameter)
# get b0 estimates for each location:
meanb0est <- modtranformed %>%
  filter(Parameter == "b_b0_Intercept") %>%
  summarize(med = median(value))
meanGP0est <- modtranformed %>%
  filter(Parameter == "b_gpP_Intercept") %>%
  summarize(med = median(value))
modtranformed %>%
  filter(grepl("r_Location__b0", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(med = median(value) + meanb0est$med + meanGP0est$med)
round(b0, 2)
# str(modtranformed)
# look at location-specific A estimates:
estA <- modtranformed %>% 
  filter(Parameter == c("b_A_Intercept"), 
         Iteration > 500) 
modtranformed %>%
  filter(grepl("r_Location__A", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(meddiff = median(value)) %>%
  mutate(med = meddiff + median(estA$value))
round(-1*A, 2)

# calculate overall avg TR for new location:
estB <- modtranformed %>% 
  filter(Parameter == c("b_B_Intercept"),
         Iteration > 500) # remove burn-in
estM <- modtranformed %>% 
  filter(Parameter == c("b_M_Intercept"),
         Iteration > 500) # remove burn-in
DATmatrix = matrix(DAT, nrow = nrow(estA), ncol = length(DAT), byrow = T)
tmp = apply(DATmatrix, 2, function(x) x - estM$value) # nIter x nDAT
beta1 = apply(tmp, 2, function(x) -1 * estA$value * (1 - 1 / (1 + exp(-1 * estB$value * x ))))
TR = 100 * (1 - exp(beta1))

TRdf <- data.frame(TR = round(apply(TR, 2, median), 1), 
           lowerCI = round(apply(TR, 2, quantile, 0.025), 1), 
          upperCI = round(apply(TR, 2, quantile, 0.975), 1),
          DAT = DAT)
ggplot(TRdf) +
  geom_line(aes(DAT, TR), size = 1.2) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI), alpha = 0.3) +
  geom_line(data = simdata1, aes(DAT, TR), color = "purple", size = 1.2) +
  labs(title = "Predicted and true TR curve")

```
```{r preds6}

## predict moth counts for each location.
preds = posterior_predict(fit6)
str(preds) # nIter x nrow(simdata)
predsDF <- bind_cols(simdata6,
                     preds = apply(preds, 2, median),
                     lowerCI = apply(preds, 2, quantile, 0.025),
                     upperCI = apply(preds, 2, quantile, 0.975))
ggplot(predsDF) +
  geom_line(aes(DAT, preds, color = Treatment)) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI, fill = Treatment), alpha = 0.3) +
  geom_point(data = simdata6, aes(DAT, nYSB, color = Treatment)) +
  facet_wrap(~Location, scales = "free_y")
#
```

## Model our case study data

Too many divergent warnings with model with intercept, A, B, M RE, with GP (DIFF for each loc). (M was multimodal)

Use constant B, M instead.

```{r fitR, eval =F}
# https://bookdown.org/content/4857/adventures-in-covariance.html#continuous-categories-and-the-gaussian-process
prior6 <- prior(normal(0,2), nlpar = "b0") +
  prior(normal(0, 2), lb = 0, nlpar = "A") + 
  prior(normal(0, 0.5), lb = 0, nlpar = "B") +
  prior(uniform(0, 120), lb = 0, ub = 120, nlpar = "M") +
  prior(cauchy(0,1), lb = 0, class = "shape") +
  prior(cauchy(0,50), lb = 0, class = "lscale", nlpar = "gpP") +
  prior(cauchy(0,10), lb = 0, class = "sdgp", nlpar = "gpP")


datR %<>%
  mutate(numTrt = ifelse(Treatment == "CGP", 0, 1))

fitR <- brm(bf(nYSB ~  b0 - (A * (1 - 1 / (1 + exp(-1 * B * (DAT - M) )))) * numTrt +
              gpP + 
              log(DaysOfCatch), 
              b0 ~ (1|Location),
              gpP ~ 1 + gp(DAT, by = Location), # GP absorbs the intercept
              A  ~ 1 + (1|Location),
              B + M ~ 1,
              nl = TRUE), 
              control = list(adapt_delta = 0.9),
              data = datR, 
              family = negbinomial,
              prior = prior6,
              warmup = 1000, iter = 2000, 
              seed = 5000,
              chains = nCores, cores = nCores) 
saveRDS(fitR, "output/fitR.RDS")
```
```{r summaryR}
fitR <- readRDS("output/fitR.RDS")
summary(fitR)
# plot(fitR)
# conditional_effects(fitR, "DAT:numTrt")
# plot(conditional_effects(fitR, "DAT:numTrt"), points = T)
# # 

```

```{r outputR}

# CALC TR
modtranformed <- ggs(fitR) # transform MCMC output into a table
unique(modtranformed$Parameter)
# get b0 estimates for each location:
meanb0est <- modtranformed %>%
  filter(Parameter == "b_b0_Intercept") %>%
  summarize(med = median(value))
meanGP0est <- modtranformed %>%
  filter(Parameter == "b_gpP_Intercept") %>%
  summarize(med = median(value))
modtranformed %>%
  filter(grepl("r_Location__b0", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(med = median(value) + meanb0est$med + meanGP0est$med)

# str(modtranformed)
# look at location-specific A estimates:
estA <- modtranformed %>% 
  filter(Parameter == c("b_A_Intercept"), 
         Iteration > 500) 
modtranformed %>%
  filter(grepl("r_Location__A", Parameter)) %>%
  group_by(Parameter) %>%
  summarize(meddiff = median(value)) %>%
  mutate(med = meddiff + median(estA$value))

# calculate overall avg TR for new location:
estB <- modtranformed %>% 
  filter(Parameter == c("b_B_Intercept"),
         Iteration > 500) # remove burn-in
estM <- modtranformed %>% 
  filter(Parameter == c("b_M_Intercept"),
         Iteration > 500) # remove burn-in
DAT = 1:150
DATmatrix = matrix(DAT, nrow = nrow(estA), ncol = length(DAT), byrow = T)
tmp = apply(DATmatrix, 2, function(x) x - estM$value) # nIter x nDAT
beta1 = apply(tmp, 2, function(x) -1 * estA$value * (1 - 1 / (1 + exp(-1 * estB$value * x ))))
TR = 100 * (1 - exp(beta1))

TRdf <- data.frame(TR = round(apply(TR, 2, median), 1), 
           lowerCI = round(apply(TR, 2, quantile, 0.025), 1), 
          upperCI = round(apply(TR, 2, quantile, 0.975), 1),
          DAT = DAT)
ggplot(TRdf) +
  geom_line(aes(DAT, TR), size = 1.2) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI), alpha = 0.3) +
  labs(title = "Predicted TR curve") +
  ylim(0, 100)

```
```{r predsR}

## predict moth counts for each location.
# preds = posterior_predict(fitR, newdata = datR %>% mutate(DaysOfCatch = 1))
predsbrm <- fitted(fitR,
                   scale = "linear",
                   newdata = datR  %>% mutate(DaysOfCatch = 1))
# str(predsbrm)
predsDF <- bind_cols(datR,
                     preds = predsbrm[,"Estimate"],
                     lowerCI = predsbrm[, "Q2.5"],
                     upperCI = predsbrm[, "Q97.5"])
# str(preds) # nIter x nrow(simdata)
# predsDF <- bind_cols(datR,
#                      preds = apply(preds, 2, median),
#                      lowerCI = apply(preds, 2, quantile, 0.025),
#                      upperCI = apply(preds, 2, quantile, 0.975))
ggplot(predsDF) +
  geom_line(aes(DAT, preds, color = Treatment)) +
  geom_ribbon(aes(DAT, ymin = lowerCI, ymax = upperCI, fill = Treatment), alpha = 0.3) +
  geom_point(data = mean_cts, aes(DATI, log(mean_mothsperday), color = Treatment)) +
  facet_wrap(~Location, scales = "free_y")
#
```
